---
title: "Population Models"
execute:
  echo: false
---

## Setup

```{r}
#| label: setup
#| include: false

library(here)
source(here("Scripts", "DataAppendixScripts", "helpers.R"))
source(here("Scripts", "DataAppendixScripts", "models.R"))
```

## Population model — Meta-analytic (Lin & Yao, 2022)

The primary population model uses factor loadings derived from the
meta-analysis by Lin and Yao (2022). Four factors: Psychological (6 items),
Physical (7 items), Social (3 items), Environment (8 items). Interfactor
correlations are set at the upper bound (0.3) reported in the meta-analysis
(range: 0.08–0.30). Residual variances follow the formula 1 − λ², except
for Q3 and Q4 which share a correlated residual (0.3).

```{r}
#| label: pop-model

cat(popModel)
```

## Naive model — Minimally acceptable scenario

All factor loadings set to 0.5 (minimum acceptable for CFA), residual
variances = 0.75 (i.e., 1 − 0.5²), and small interfactor correlations (0.3).
This represents a **pessimistic** scenario.

```{r}
#| label: naive-model

cat(naiveModel)
```

## Optimistic model — Best-case scenario

All factor loadings set to 0.7 (strong loadings), residual variances = 0.51
(1 − 0.7²), and moderate interfactor correlations (0.5). This represents an
**optimistic** scenario.

```{r}
#| label: opt-model

cat(optModel)
```

## Analysis model — Free parameters

This is the model the researcher fits to data. All parameters are freely
estimated (no fixed values), matching the structure of the population models.

```{r}
#| label: analysis-model

cat(analyzeModel)
```

## Comparison table

```{r}
#| label: tbl-pop-comparison
#| tbl-cap: "Summary of population model variants used for sensitivity analysis"

pop_summary <- data.frame(
  Model = c("Meta-analytic (popModel)",
            "Naive (naiveModel)",
            "Optimistic (optModel)"),
  `Loading Range` = c("0.37–0.94", "0.50", "0.70"),
  `Mean Loading` = c("0.74", "0.50", "0.70"),
  `Factor Correlations` = c("0.30", "0.30", "0.50"),
  `Residual Range` = c("0.12–0.86", "0.75", "0.51"),
  `Q3–Q4 Covariance` = c("0.30", "0.30", "0.50"),
  check.names = FALSE
)

kable(pop_summary, align = "lccccc")
```
