---
title: "Analytical Power Analysis"
execute:
  echo: false
---

## Setup

```{r}
#| label: setup
#| include: false

library(here)
source(here("Scripts", "DataAppendixScripts", "helpers.R"))
source(here("Scripts", "DataAppendixScripts", "models.R"))
```


## Computing model degrees of freedom

The degrees of freedom are critical for the analytical approach because
power is computed from the central and non-central χ² distributions,
both parameterized by df.

```{r}
#| label: compute-df

# Compute df directly with semPower (no data needed)
df_model <- semPower::semPower.getDf(analyzemodel)

cat("Model degrees of freedom (via semPower):", df_model, "\n")
```

## A priori power analysis: Determine N

### Model-free approach (RMSEA-based)

This chunk uses `semPower::semPower.aPriori()` to compute the minimum sample size required for an a priori power analysis based on RMSEA, given the target \(\alpha\), desired power, and model degrees of freedom [@moshagen2023].

The output returns the required number of observations and the associated quantities reported by `semPower`, such as \(F_0\), RMSEA, Mc, the critical chi-square value, and the noncentrality parameter.

```{r}
#| label: apriori-rmsea

# Analytical a priori power analysis for the global chi-square test.
# The goal is to determine the minimum N required to detect
# a population misfit corresponding to RMSEA = 0.05.

ap_rmsea <- semPower::semPower.aPriori(
  effect         = 0.05,      # RMSEA under H1
  effect.measure = "RMSEA",   # effect is expressed as RMSEA
  alpha          = ALPHA,     # Type I error rate
  power          = POWER,     # desired statistical power = 1 - beta
  df             = df_model,  # model degrees of freedom
  p              = P          # only relevant for indices such as GFI/AGFI
)

# Shows effect size, df, required N, critical chi-square,
# noncentrality parameter, and implied alpha/beta results.
summary(ap_rmsea)
```

### Model-based approach (with lavaan syntax)

This chunk runs an **a priori analytical power analysis** with `semPower.powerLav()`. The function uses the population model (`modelPop`) and the null analysis model (`modelH0`) to determine the minimum sample size required to detect their discrepancy with the target alpha and power [@moshagen2023; @jobst2023].

Because no `modelH1` is specified, the comparison is made against the **saturated model**, so the test targets global model misfit [@moshagen2023; @jobst2023].

```{r}
#| label: apriori-popmodel

# A priori power analysis:
# computes the minimum N needed to detect the discrepancy
# between the population model and the null analysis model.

ap_popmodel <- semPower.powerLav(
  type = "a-priori",   # sample size planning
  alpha = ALPHA,       # Type I error rate
  power = POWER,       # target power
  modelPop = popmodel, # population-generating model
  modelH0 = analyzemodel, # null/analysis model
  p = P                 # only relevant for indices such as GFI/AGFI
)

# Shows effect size, df, required N, critical chi-square,
# noncentrality parameter, and implied alpha/beta results.
summary(ap_popmodel)
```

## Power curves

### Model-free approach

This chunk generates a model-free power curve showing how statistical power changes as a function of sample size \(N\) for a fixed effect defined as \(\mathrm{RMSEA} = 0.05\), given the chosen significance level \(\alpha\) and the model degrees of freedom \(df\) [@moshagen2023; @jobst2023].

In `semPower`, this is an analytical power plot for a noncentrality-based effect measure, so the curve summarizes the expected power across a range of sample sizes while holding the effect size and test conditions constant [@moshagen2023; @jobst2023].

```{r}
#| label: fig-power-rmsea
#| fig-cap: "Power as a function of sample size for RMSEA >= 0.05"
#| fig-width: 8
#| fig-height: 5

# Plots statistical power against sample size N.
semPower::semPower.powerPlot.byN(
  effect         = 0.05, # effect = 0.05 sets the target misfit in RMSEA units.
  effect.measure = "RMSEA",
  alpha          = ALPHA, # alpha = ALPHA sets the Type I error rate.
  df             = df_model, # df = df_model sets the model degrees of freedom.
  power.min      = 0.05,
  power.max      = 0.99 # power.min and power.max define the y-axis range of the plotted curve.
)
```

### Model-based approach

This chunk produces a power curve for the model-based analysis stored in `ap_popmodel`, displaying how statistical power changes as a function of sample size \(N\) for the effect defined in that object [@moshagen2023; @jobst2023]. In `semPower`, `semPower.powerPlot()` is used to visualize the power trajectory implied by a previously estimated power-analysis object, which is useful for identifying the sample-size region in which power becomes adequate [@moshagen2023; @jobst2023].

```{r}
#| label: fig-power-popmodel
#| fig-cap: "Power as a function of sample size for the population model (popModel)"
#| fig-width: 8
#| fig-height: 5

# Plots the power curve based on the previously computed object `ap_popmodel`.
# x-axis: sample size (N).
# y-axis: statistical power.

semPower.powerPlot(ap_popmodel)
```

## Summary table

```{r}
#| label: tbl-analytical-summary
#| tbl-cap: "Summary of analytical power analysis results"

# Helper to safely extract a single numeric value from each result object
safe_num <- function(x) if (length(x) == 0 || is.null(x)) NA else as.numeric(x)[1]

analytical_summary <- data.frame(
    Method = c(
        "Model-free a priori (RMSEA ≥ .05)",
        "Model-based a priori (Population Model)"
    ),
    `N Required` = c(
        safe_num(ap_rmsea$requiredN),
        safe_num(ap_popmodel$requiredN)
    ),
    `Power` = c(
        round(safe_num(ap_rmsea$impliedPower), 4),
        round(safe_num(ap_popmodel$impliedPower), 4)
    ),
    `Implied RMSEA` = c(
        round(safe_num(ap_rmsea$rmsea), 4),
        round(safe_num(ap_popmodel$rmsea), 4)
    ),
    df = rep(safe_num(df_model), 2),
    check.names = FALSE
)

# Render the summary table with simple column alignment for reporting
knitr::kable(analytical_summary, align = "lccccc")
```
