---
title: "Analytical Power Analysis"
execute:
  echo: false
  warning: false
  message: false
---

## Setup

```{r}
#| label: setup
#| include: false

# For CFA population models
source(here::here("Scripts", "DataAppendixScripts", "models.R"))

# For DRY (Don't repeat yourself) parameters
source(here::here("Scripts", "DataAppendixScripts", "helpers.R"))
```

## Computing model degrees of freedom

Before any analytical power calculation can be performed, the model
degrees of freedom ($df$) must be established. This quantity is the
difference between the number of unique variances and covariances
in the observed data and the number of freely estimated parameters.
It enters the non-central $\chi^2$ distribution directly and is
therefore required for both the a priori and the power-curve analyses
that follow [@jak2021; @moshagen2024].

```{r}
#| label: compute-df

# Compute df for standardized latent variables
# 24 observed variables: p*(p+1)/2 = 300 non-redundant sample moments
# analyzemodel: 24 loadings + 6 factor correlations + 1 residual covariance (Q3~~Q4) + 24 residual variances = 55 free parameters
# Model degrees of freedom: df = 300 - 55 = 245
df_model <- semPower.getDf(analyzemodel)-4
# Minus 4 because the 4 residual variances are fixed to 1 in analyzemodel, 
# and in the foward analysis we use std.lv = TRUE (standardized latent variables). 

cat("Model degrees of freedom:", df_model, "\n")
```

## A priori power analysis: Determine N

### Model-free approach (RMSEA-based)

The model-free strategy determines the required sample size from four
ingredients — the target RMSEA, the significance level, the desired
power, and the model degrees of freedom — without requiring an
explicitly specified population model [@maccallum1996; @jak2021].
Because RMSEA expresses misfit per degree of freedom, it translates
directly into the noncentrality parameter $\lambda = (N-1)\,df\,\text{RMSEA}^2$,
which locates the alternative $\chi^2$ distribution relative to the
null [@maccallum1996].

The output below returns the minimum $N$ needed to detect
$\text{RMSEA} = 0.05$ under the chosen $\alpha$ and power, together
with equivalent representations of the same effect in terms of $F_0$,
$Mc$, $GFI$, and $AGFI$ [@moshagen2024].

```{r}
#| label: fig-apriori-rmsea
#| fig-cap: "A priori power analysis for exact fit with RMSEA-based effect size"

# Analytical a priori power analysis for the global chi-square test.
# The goal is to determine the minimum N required to detect
# a population misfit corresponding to RMSEA = 0.05.

ap_rmsea <- semPower::semPower.aPriori(
  effect         = 0.05,      # RMSEA under H1
  effect.measure = "RMSEA",   # effect is expressed as RMSEA
  alpha          = ALPHA,     # Type I error rate defined in helpers.R
  power          = POWER,     # desired statistical power defined in helpers.R
  df             = df_model,  # model degrees of freedom calculated in "compute-df" chunk
  p              = P          # number of observed indicators defined in helpers.R
)

# Shows effect size, df, required N, critical chi-square,
# noncentrality parameter, and implied alpha/beta results.
summary(ap_rmsea)
```

### Model-based approach (with lavaan syntax)

When the substantive hypothesis is grounded in an explicit population
model, the effect no longer needs to be set by convention: it is
implied by the discrepancy between the population covariance structure
and the restrictions imposed by the analysis model [@wang2023; @feng2023].
This makes the sample size estimate more directly tied to the CFA
under study, at the cost of requiring a fully specified population
model as input [@buchberger2024; @moshagen2024].

Here, `semPower.powerLav()` fits the analysis model (`analyzemodel`) to the
covariance matrix implied by the misspecified population model (`h1model`)
and translates the resulting discrepancy into $F_0$ and then into $\lambda = (N-1)F_0$.
Because `modelH1` is not specified, the comparison is
against the saturated model, evaluating overall model misfit ($df = 245$).

```{r}
#| label: fig-apriori-popmodel
#| fig-cap: "A priori power analysis Model-based approach (h1model vs analyzemodel)"

# A priori power analysis:
# computes the minimum N needed to detect the omission of the three
# cross-loadings in analyzemodel when data arise from h1model.

ap_popmodel <- semPower::semPower.powerLav(
  type       = "a-priori", # Type of power analysis
  alpha      = ALPHA,      # Type I error rate defined in helpers.R
  power      = POWER,      # desired statistical power defined in helpers.R
  modelPop   = h1model,    # population model defined in "models.R"
  modelH0    = analyzemodel, # analysis model defined in "models.R"
  lavOptions = list(std.lv = TRUE) # Standardize latent variables
)

# Shows effect size, df, required N, critical chi-square,
# noncentrality parameter, and implied alpha/beta results.
summary(ap_popmodel)
```

## Power curves

Power curves shift the discussion from a single required-$N$ result
to a continuous view of how detection probability changes as sample
size grows. They are useful for communicating the design in a way
that makes trade-offs between $N$, power, and tolerated misspecification
immediately visible [@wang2023; @moshagen2024].

### Model-free approach

The curve below is analytical and model-free: power is a function
of RMSEA, $df$, and $\alpha$ only. It answers the planning question
of how many observations are needed to detect a fixed degree of
global misspecification across a range of possible $N$ values
[@jak2021; @jobst2023].

```{r}
#| label: fig-power-rmsea
#| fig-cap: "Power as a function of sample size for RMSEA = 0.05"
#| fig-width: 8
#| fig-height: 5

# Plots statistical power against sample size N.
semPower::semPower.powerPlot.byN(
  effect         = 0.05,      # Target misfit in RMSEA units.
  effect.measure = "RMSEA",   # Type of effect measure
  alpha          = ALPHA,     # Type I error rate defined in helpers.R
  df             = df_model,  # Model degrees of freedom calculated in "compute-df" chunk
  power.min      = 0.05,      # Min. Power define the y-axis range
  power.max      = 0.99       # Max. Power define the y-axis range
)
```

### Model-based approach

This curve traces the same power function for the model-based effect
stored in `ap_popmodel`. Because the effect here is derived from the
population covariance structure, the curve reflects the specific
degree of misspecification implied by the substantive CFA, rather
than a generic RMSEA threshold [@feng2023; @buchberger2024].
The sample-size region in which the curve crosses the target power
level is the practical read-off from this figure [@moshagen2024].

```{r}
#| label: fig-power-popmodel
#| fig-cap: "Power as a function of sample size for the population model (popModel)"
#| fig-width: 8
#| fig-height: 5

# Plots the power curve based on the previously computed object `ap_popmodel`.
# x-axis: sample size (N).
# y-axis: statistical power.

semPower.powerPlot(ap_popmodel)
```

## Summary table

The table below brings the two analytical solutions together,
allowing a direct comparison of their sample size requirements under
the same $\alpha$ and power targets. The difference between the two
estimates reflects a general principle: once the effect is derived
from a specific population model rather than from a generic RMSEA
convention, the resulting $N$ becomes more substantively grounded
and, in this case, more conservative [@wang2023; @feng2023].

```{r}
#| label: tbl-analytical-summary
#| tbl-cap: "Summary of analytical power analysis results"

# Helper to safely extract a single numeric value from each result object
safe_num <- function(x) if (length(x) == 0 || is.null(x)) NA else as.numeric(x)[1]

# Create a summary table of the analytical power analysis results
analytical_summary <- data.frame(
    Method = c(
        "Model-free a priori (RMSEA = .05)",
        "Model-based global fit (h1model vs. saturated)"
    ),
    `N Required` = c(
        safe_num(ap_rmsea$requiredN),
        safe_num(ap_popmodel$requiredN)
    ),
    `Power` = c(
        round(safe_num(ap_rmsea$impliedPower), 4),
        round(safe_num(ap_popmodel$impliedPower), 4)
    ),
    `Implied RMSEA` = c(
        round(safe_num(ap_rmsea$rmsea), 4),
        round(safe_num(ap_popmodel$rmsea), 4)
    ),
    df = c(
        safe_num(ap_rmsea$df),
        safe_num(ap_popmodel$df)
    ),
    check.names = FALSE
)

# Render the summary table with simple column alignment
knitr::kable(analytical_summary, align = "lcccc")
```
