---
title: "Analytical Power Analysis"
execute:
  echo: false
---

## Setup

```{r}
#| label: setup
#| include: false

library(here)
source(here("Scripts", "DataAppendixScripts", "helpers.R"))
source(here("Scripts", "DataAppendixScripts", "models.R"))
```


## Computing model degrees of freedom

The degrees of freedom are critical for the analytical approach because
power is computed from the central and non-central χ² distributions,
both parameterized by df.

```{r}
#| label: compute-df

# Compute df directly with semPower (no data needed)
df_model <- semPower::semPower.getDf(analyzemodel)

cat("Model degrees of freedom (via semPower):", df_model, "\n")
```

## A priori power analysis: Determine N

### Model-free approach (RMSEA-based)

Given the model df and a target RMSEA of 0.05, how many observations are
needed for 80% power to detect misfit?

```{r}
#| label: apriori-rmsea

ap_rmsea <- semPower::semPower.aPriori(
    effect         = 0.05, # RMSEA under H1 (exact-fit model: H0: RMSEA = 0)
    effect.measure = "RMSEA",
    alpha          = ALPHA,
    power          = POWER,
    df             = df_model,
    p              = P # Just for GFI and AGFI
)

summary(ap_rmsea)
```

### Model-based approach (with lavaan syntax)

```{r}
#| label: apriori-popmodel

ap_popmodel <- semPower.powerLav(
  type = 'a-priori',
  alpha = ALPHA,
  power = POWER,
  modelPop = popmodel,
  modelH0 = analyzemodel
)

summary(ap_popmodel)
```

## Power curves

### Model-free approach

```{r}
#| label: fig-power-rmsea
#| fig-cap: "Power as a function of sample size for RMSEA ≥ 0.05"
#| fig-width: 8
#| fig-height: 5

semPower::semPower.powerPlot.byN(
    effect         = 0.05,
    effect.measure = "RMSEA",
    alpha          = ALPHA,
    df             = df_model,
    power.min      = 0.05,
    power.max      = 0.99
)
```

### Model-based approach

```{r}
#| label: fig-power-popmodel
#| fig-cap: "Power as a function of sample size for the population model (popModel)"
#| fig-width: 8
#| fig-height: 5

semPower.powerPlot(ap_popmodel)
```

## Summary table

```{r}
#| label: tbl-analytical-summary
#| tbl-cap: "Summary of analytical power analysis results"

# Helper to ensure length-1 numeric
safe_num <- function(x) if (length(x) == 0 || is.null(x)) NA else as.numeric(x)[1]

analytical_summary <- data.frame(
    Method = c(
        "Model-free a priori (RMSEA ≥ .05)",
        "Model-based a priori (Population Model)"
    ),
    `N Required` = c(
        safe_num(ap_rmsea$requiredN),
        safe_num(ap_popmodel$requiredN)
    ),
    `Power` = c(
        round(safe_num(ap_rmsea$impliedPower), 4),
        round(safe_num(ap_popmodel$impliedPower), 4)
    ),
    `Implied RMSEA` = c(
        round(safe_num(ap_rmsea$rmsea), 4),
        round(safe_num(ap_popmodel$rmsea), 4)
    ),
    `Implied Fmin` = c(
        round(safe_num(ap_rmsea$fmin), 4),
        round(safe_num(ap_popmodel$fmin), 4)
    ),
    df = rep(safe_num(df_model), 2),
    check.names = FALSE
)

knitr::kable(analytical_summary, align = "lccccc")
```
