---
title: "Analytical Power Analysis"
execute:
  echo: false
  warning: false
  message: false
---

## Setup

```{r}
#| label: setup
#| include: false

library(here)
source(here("Scripts", "DataAppendixScripts", "models.R"))
source(here("Scripts", "DataAppendixScripts", "helpers.R"))
```

## Computing model degrees of freedom

Before any analytical power calculation can be performed, the model
degrees of freedom ($df$) must be established. This quantity is the
difference between the number of unique variances and covariances
in the observed data and the number of freely estimated parameters.
It enters the non-central $\chi^2$ distribution directly and is
therefore required for both the a priori and the power-curve analyses
that follow [@jak2021; @moshagen2024].

```{r}
#| label: compute-df

# Compute df directly with semPower (no data needed)
df_model <- semPower::semPower.getDf(analyzemodel)

cat("Model degrees of freedom (via semPower):", df_model, "\n")
```

## A priori power analysis: Determine N

### Model-free approach (RMSEA-based)

The model-free strategy determines the required sample size from four
ingredients — the target RMSEA, the significance level, the desired
power, and the model degrees of freedom — without requiring an
explicitly specified population model [@maccallum1996; @jak2021].
Because RMSEA expresses misfit per degree of freedom, it translates
directly into the noncentrality parameter $\lambda = (N-1)\,df\,\text{RMSEA}^2$,
which locates the alternative $\chi^2$ distribution relative to the
null [@maccallum1996].

The output below returns the minimum $N$ needed to detect
$\text{RMSEA} = 0.05$ under the chosen $\alpha$ and power, together
with equivalent representations of the same effect in terms of $F_0$,
$Mc$, $GFI$, and $AGFI$ [@moshagen2024].

```{r}
#| label: fig-apriori-rmsea
#| fig-cap: "A priori power analysis for exact fit with RMSEA-based effect size"

# Analytical a priori power analysis for the global chi-square test.
# The goal is to determine the minimum N required to detect
# a population misfit corresponding to RMSEA = 0.05.

ap_rmsea <- semPower::semPower.aPriori(
  effect         = 0.05,      # RMSEA under H1
  effect.measure = "RMSEA",   # effect is expressed as RMSEA
  alpha          = ALPHA,     # Type I error rate
  power          = POWER,     # desired statistical power = 1 - beta
  df             = df_model,  # model degrees of freedom
  p              = P          # only relevant for indices such as GFI/AGFI
)

# Shows effect size, df, required N, critical chi-square,
# noncentrality parameter, and implied alpha/beta results.
summary(ap_rmsea)
```

### Model-based approach (with lavaan syntax)

When the substantive hypothesis is grounded in an explicit population
model, the effect no longer needs to be set by convention: it is
implied by the discrepancy between the population covariance structure
and the restrictions imposed by the analysis model [@wang2023; @feng2023].
This makes the sample size estimate more directly tied to the CFA
under study, at the cost of requiring a fully specified population
model as input [@buchberger2024; @moshagen2024].

Here, `semPower.powerLav()` fits the analysis model to the
population-implied covariance matrix and translates the resulting
discrepancy into $F_0$ and then into $\lambda = (N-1)F_0$. Because
`modelH1` is not specified, the comparison is against the saturated
model, so this remains a global test of fit rather than a test of
any individual restriction [@jobst2023; @moshagen2024].

```{r}
#| label: fig-apriori-popmodel
#| fig-cap: "A priori power analysis Model-based approach (popModel)"

# A priori power analysis:
# computes the minimum N needed to detect the discrepancy
# between the population model and the null analysis model.

ap_popmodel <- semPower.powerLav(
  type = "a-priori",       # sample size planning
  alpha = ALPHA,           # Type I error rate
  power = POWER,           # target power
  modelPop = popmodel,     # population-generating model
  modelH0 = analyzemodel,  # null/analysis model
  p = P                    # only relevant for indices such as GFI/AGFI
)

# Shows effect size, df, required N, critical chi-square,
# noncentrality parameter, and implied alpha/beta results.
summary(ap_popmodel)
```

## Power curves

Power curves shift the discussion from a single required-$N$ result
to a continuous view of how detection probability changes as sample
size grows. They are useful for communicating the design in a way
that makes trade-offs between $N$, power, and tolerated misspecification
immediately visible [@wang2023; @moshagen2024].

### Model-free approach

The curve below is analytical and model-free: power is a function
of RMSEA, $df$, and $\alpha$ only. It answers the planning question
of how many observations are needed to detect a fixed degree of
global misspecification across a range of possible $N$ values
[@jak2021; @jobst2023].

```{r}
#| label: fig-power-rmsea
#| fig-cap: "Power as a function of sample size for RMSEA = 0.05"
#| fig-width: 8
#| fig-height: 5

# Plots statistical power against sample size N.
semPower::semPower.powerPlot.byN(
  effect         = 0.05,      # effect = 0.05 sets the target misfit in RMSEA units.
  effect.measure = "RMSEA",
  alpha          = ALPHA,     # alpha = ALPHA sets the Type I error rate.
  df             = df_model,  # df = df_model sets the model degrees of freedom.
  power.min      = 0.05,
  power.max      = 0.99       # power.min and power.max define the y-axis range of curve.
)
```

### Model-based approach

This curve traces the same power function for the model-based effect
stored in `ap_popmodel`. Because the effect here is derived from the
population covariance structure, the curve reflects the specific
degree of misspecification implied by the substantive CFA, rather
than a generic RMSEA threshold [@feng2023; @buchberger2024].
The sample-size region in which the curve crosses the target power
level is the practical read-off from this figure [@moshagen2024].

```{r}
#| label: fig-power-popmodel
#| fig-cap: "Power as a function of sample size for the population model (popModel)"
#| fig-width: 8
#| fig-height: 5

# Plots the power curve based on the previously computed object `ap_popmodel`.
# x-axis: sample size (N).
# y-axis: statistical power.

semPower.powerPlot(ap_popmodel)
```

## Summary table

The table below brings the two analytical solutions together,
allowing a direct comparison of their sample size requirements under
the same $\alpha$ and power targets. The difference between the two
estimates reflects a general principle: once the effect is derived
from a specific population model rather than from a generic RMSEA
convention, the resulting $N$ becomes more substantively grounded
and, in this case, more conservative [@wang2023; @feng2023].

```{r}
#| label: tbl-analytical-summary
#| tbl-cap: "Summary of analytical power analysis results"

# Helper to safely extract a single numeric value from each result object
safe_num <- function(x) if (length(x) == 0 || is.null(x)) NA else as.numeric(x)[1]

analytical_summary <- data.frame(
    Method = c(
        "Model-free a priori (RMSEA = .05)",
        "Model-based a priori (Population Model)"
    ),
    `N Required` = c(
        safe_num(ap_rmsea$requiredN),
        safe_num(ap_popmodel$requiredN)
    ),
    `Power` = c(
        round(safe_num(ap_rmsea$impliedPower), 4),
        round(safe_num(ap_popmodel$impliedPower), 4)
    ),
    `Implied RMSEA` = c(
        round(safe_num(ap_rmsea$rmsea), 4),
        round(safe_num(ap_popmodel$rmsea), 4)
    ),
    df = rep(safe_num(df_model), 2),
    check.names = FALSE
)

# Render the summary table with simple column alignment for reporting
knitr::kable(analytical_summary, align = "lccccc")
```
