---
title: "Analytical Power Analysis"
execute:
  echo: false
---

## Setup

```{r}
#| label: setup
#| include: false

library(here)
source(here("Scripts", "DataAppendixScripts", "helpers.R"))
source(here("Scripts", "DataAppendixScripts", "models.R"))
```


## Computing model degrees of freedom

The degrees of freedom are critical for the analytical approach because
power is computed from the central and non-central χ² distributions,
both parameterized by df.

```{r}
#| label: compute-df

# Method 1: Generate data from the population model and fit the analysis model
sim_data <- lavaan::simulateData(popModel, sample.nobs = 500)
fit_temp <- lavaan::cfa(analyzeModel, data = sim_data, std.lv = TRUE)
df_model <- lavaan::fitMeasures(fit_temp, "df")

cat("Model degrees of freedom (via lavaan):", df_model, "\n")

# Method 2: Compute df directly with semPower (no data needed)
# p = 24 indicators, q = free parameters
# df = p(p+1)/2 - q
p <- 24
n_unique <- p * (p + 1) / 2  # 300 unique elements in covariance matrix
cat("Unique covariance elements:", n_unique, "\n")
cat("Model df:", df_model, "\n")
```

## A priori power analysis: Determine N

### Model-free approach (RMSEA-based) with semPower

Given the model df and a target RMSEA of 0.05, how many observations are
needed for 80% power to detect misfit?

```{r}
#| label: apriori-rmsea

ap_rmsea <- semPower::semPower.aPriori(
  effect         = 0.05,
  effect.measure = "RMSEA",
  alpha          = ALPHA,
  power          = POWER_TARGET,
  df             = df_model
)

summary(ap_rmsea)
```

### A priori N with semTools

Cross-validation: semTools uses the same underlying χ² approach and should
yield concordant results.

```{r}
#| label: apriori-semtools

n_semtools <- semTools::findRMSEAsamplesize(
  rmsea0 = 0.05,    # RMSEA under H1 (the effect to detect)
  rmseaA = 0,       # RMSEA under H0 (perfect fit)
  df     = df_model,
  power  = POWER_TARGET,
  alpha  = ALPHA
)

cat("Required N (semTools):", n_semtools, "\n")
```

## Post hoc power analysis: Achieved power with N = 1047

### With semPower

```{r}
#| label: posthoc-sempower

ph_sempower <- semPower::semPower.postHoc(
  effect         = 0.05,
  effect.measure = "RMSEA",
  alpha          = ALPHA,
  N              = N_EMPIRICAL,
  df             = df_model
)

summary(ph_sempower)
```

### With semTools

```{r}
#| label: posthoc-semtools

power_semtools <- semTools::findRMSEApower(
  rmsea0 = 0.05,
  rmseaA = 0,
  df     = df_model,
  n      = N_EMPIRICAL,
  alpha  = ALPHA
)

cat("Achieved power (semTools):", round(power_semtools, 4), "\n")
```

## Power curves

### Power by sample size

```{r}
#| label: fig-power-by-n
#| fig-cap: "Power as a function of sample size for RMSEA ≥ 0.05"
#| fig-width: 8
#| fig-height: 5

semPower::semPower.powerPlot.byN(
  effect         = 0.05,
  effect.measure = "RMSEA",
  alpha          = ALPHA,
  df             = df_model,
  power.min      = 0.05,
  power.max      = 0.99
)
```

### Power by effect size

```{r}
#| label: fig-power-by-effect
#| fig-cap: "Power as a function of RMSEA magnitude with N = 1047"
#| fig-width: 8
#| fig-height: 5

semPower::semPower.powerPlot.byEffect(
  effect.measure = "RMSEA",
  alpha          = ALPHA,
  N              = N_EMPIRICAL,
  df             = df_model,
  effect.min     = 0.001,
  effect.max     = 0.10
)
```

## Summary table

```{r}
#| label: tbl-analytical-summary
#| tbl-cap: "Summary of analytical power analysis results"

# Helper to ensure length-1 numeric
safe_num <- function(x) if(length(x) == 0 || is.null(x)) NA else as.numeric(x)[1]

analytical_summary <- data.frame(
  Method = c(
    "semPower a priori (RMSEA ≥ .05)",
    "semTools a priori (RMSEA ≥ .05)",
    "semPower post hoc (N = 1047)",
    "semTools post hoc (N = 1047)"
  ),
  `N Required` = c(
    safe_num(ap_rmsea$requiredN),
    safe_num(n_semtools),
    N_EMPIRICAL,
    N_EMPIRICAL
  ),
  `Power` = c(
    round(safe_num(ap_rmsea$power), 4),
    POWER_TARGET,
    round(safe_num(ph_sempower$power), 4),
    round(safe_num(power_semtools), 4)
  ),
  df = rep(safe_num(df_model), 4),
  check.names = FALSE
)

kable(analytical_summary, align = "lccc")
```
