---
title: "Simulation — Ideal Conditions"
execute:
  echo: false
---

## Setup

```{r}
#| label: setup
#| include: false

library(here)
source(here("Scripts", "DataAppendixScripts", "helpers.R"))
source(here("Scripts", "DataAppendixScripts", "models.R"))
```

## Main misspecification simulation

Coul be create an trivial (or harder) artificial misspecification. Just one rep by each N: in your case you should be use more reps, but for the sake of time we will use just one.

```{r}
#| label: sim-ideal-popmodel
#| cache: true

popmodel_out <- sim(
    NULL,
    n = SEQ,
    analyzemodel,
    generate = popmodel,
    std.lv = TRUE,
    lavaanfun = "cfa",
    seed = SEED,
    multicore = TRUE
)
```


```{r}
#| label: sim-ideal-h1model
#| cache: true

h1model_out <- sim(
    NULL,
    n = SEQ,
    analyzemodel,
    generate = h1model,
    std.lv = TRUE,
    lavaanfun = "cfa",
    seed = SEED,
    multicore = TRUE
)
```

## Main misspecification outputs

### Power curves for fit indices

```{r}
#| label: fig-power-curve-ideal
#| fig-cap: "Power curves for fit indices under ideal conditions."
#| fig-width: 8
#| fig-height: 5

plotPowerFit(
    h1model_out,
    popmodel_out,
    alpha = ALPHA,
    usedFit = FITS
)
```


```{r}
#| label: fig-fit-curve-ideal
#| fig-cap: "Power curves for fit indices under ideal conditions without logistic smoothing."
#| fig-width: 8
#| fig-height: 5

plotPowerFit(
    h1model_out,
    popmodel_out,
    alpha = ALPHA,
    usedFit = FITS,
    logistic = FALSE
)
```

### Power cutoff values for fit indices

You extract the cutoff values for fit indices at a given alpha level and sample size to determine the threshold for acceptable fit under the null model. This allows you to see how stringent the fit criteria are at different sample sizes.

```{r}
#| label: tbl-power-cutoff-ideal
#| tbl-cap: "Power for fit index cutoffs under ideal conditions at N = 231."

getPowerFit(
    h1model_out,
    nullObject = popmodel_out,
    alpha = ALPHA,
    nVal = 231,
    usedFit = FITS
)
```

```{r}
#| label: tbl-fit-cutoff-ideal
#| tbl-cap: "Fit index cutoff values under ideal conditions at N = 231."

getCutoff(
    popmodel_out,
    alpha = ALPHA,
    nVal = 231,
    usedFit = FITS
)
```

### Tradidional cutoff values for fit indices


```{r}
#| label: fig-fit-curve-traditional
#| fig-cap: "Power curves using traditional fit index cutoffs under ideal conditions."
#| fig-width: 8
#| fig-height: 5

plotPowerFit(
    h1model_out,
    popmodel_out,
    cutoff = RULE_OF_THUMB,
    alpha = ALPHA,
    usedFit = FITS,
    logistic = FALSE
)
```


```{r}
#| label: tbl-power-cutoff-traditional
#| tbl-cap: "Power for traditional fit index cutoffs under ideal conditions at N = 231."

getPowerFit(
    h1model_out,
    RULE_OF_THUMB,
    alpha = ALPHA,
    nVal = 231,
    usedFit = FITS
)
```

```{r}
#| label: tbl-fit-cutoff-traditional
#| tbl-cap: "Traditional fit index cutoff values under ideal conditions."

summaryFit(h1model_out, alpha = ALPHA, usedFit = FITS)
```

### Target parameter estimates


```{r}
#| label: tbl-param-estimate-ideal
#| tbl-cap: "Summary of parameter estimates under ideal conditions."

summaryParam(h1model_out, alpha = ALPHA, digits = 3, detail = TRUE)
```

```{r}
#| label: tbl-power-n-ideal
#| tbl-cap: "Sample size needed for 80% power under ideal conditions."

pow_h1 <- getPower(h1model_out, alpha = ALPHA)
findPower(pow_h1, "N", 0.80)
```

### Recovering the population parameters

Recovery of population parameters is not best strategy in CFA power analyses. The idea is to determine the sample size needed to accurately estimate the population parameters (e.g., factor loadings <> 0) with a certain level of precision. This can be done by simulating data under different models and evaluating how well the analysis model recovers the population parameters. 

But this strategy its not pratical because uselly we don't need an large sample size to recover the loadings > 0, but rather to detect misspecifications (e.g., cross-loadings) that are not present in the population model. Therefore, we can use the power to detect misspecifications as a more relevant criterion for determining the required sample size.

Naive model: tau-equivalent loadings (uniform λ per domain) with same reliability as the population model. This is a common but often unrealistic assumption in CFA power analyses.

```{r}
#| label: sim-ideal-naivemodel
#| cache: true

naivemodel_out <- sim(
    NULL,
    n = SEQ,
    analyzemodel,
    generate = naivemodel,
    std.lv = TRUE,
    lavaanfun = "cfa",
    seed = SEED,
    multicore = TRUE
)
```

Maybe the better strategy is recovering N from the lowest population loadings.

```{r}
#| label: tbl-power-n-naive
#| tbl-cap: "Sample size needed for 80% power under naive model."

pow_naive <- getPower(naivemodel_out, alpha = ALPHA)
findPower(pow_naive, "N", 0.80)
```

Optimal model: tau-equivalent loadings (uniform λ per domain) with same reliability as the population model, but with higher loadings to match the population reliability. This represents an idealized scenario where the analysis model is correctly specified in terms of reliability, but still assumes tau-equivalence.

```{r}
#| label: sim-ideal-optmodel
#| cache: true

optmodel_out <- sim(
    NULL,
    n = SEQ,
    analyzemodel,
    generate = optmodel,
    std.lv = TRUE,
    lavaanfun = "cfa",
    seed = SEED,
    multicore = TRUE
)
```

Maybe the better strategy is recovering N from the lowest population loadings.

```{r}
#| label: tbl-power-n-optimal
#| tbl-cap: "Sample size needed for 80% power under optimal model."

pow_opt <- getPower(optmodel_out, alpha = ALPHA)
findPower(pow_opt, "N", 0.80)
```

### Misspecification simulation with different population parameters

#### Naive model

- Power curves for fit indices

```{r}
#| label: fig-power-curve-naive
#| fig-cap: "Power curves for fit indices using the naive population model."
#| fig-width: 8
#| fig-height: 5

plotPowerFit(
    h1model_out,
    naivemodel_out,
    alpha = ALPHA,
    usedFit = FITS
)
```


```{r}
#| label: fig-fit-curve-naive
#| fig-cap: "Power curves for fit indices using the naive population model without logistic smoothing."
#| fig-width: 8
#| fig-height: 5

plotPowerFit(
    h1model_out,
    naivemodel_out,
    alpha = ALPHA,
    usedFit = FITS,
    logistic = FALSE
)
```

- Power cutoff values for fit indices

```{r}
#| label: tbl-power-cutoff-naive
#| tbl-cap: "Power for fit index cutoffs using the naive population model at N = 330."

getPowerFit(
    h1model_out,
    nullObject = naivemodel_out,
    alpha = ALPHA,
    nVal = 330,
    usedFit = FITS
)
```

```{r}
#| label: tbl-fit-cutoff-naive
#| tbl-cap: "Fit index cutoff values using the naive population model at N = 330."

getCutoff(
    naivemodel_out,
    alpha = ALPHA,
    nVal = 330,
    usedFit = FITS
)
```

#### Optimal model

- Power curves for fit indices

```{r}
#| label: fig-power-curve-opt
#| fig-cap: "Power curves for fit indices using the optimal population model."
#| fig-width: 8
#| fig-height: 5

plotPowerFit(
    h1model_out,
    optmodel_out,
    alpha = ALPHA,
    usedFit = FITS
)
```


```{r}
#| label: fig-fit-curve-opt
#| fig-cap: "Power curves for fit indices using the optimal population model without logistic smoothing."
#| fig-width: 8
#| fig-height: 5

plotPowerFit(
    h1model_out,
    optmodel_out,
    alpha = ALPHA,
    usedFit = FITS,
    logistic = FALSE
)
```

- Power cutoff values for fit indices

```{r}
#| label: tbl-power-cutoff-opt
#| tbl-cap: "Power for fit index cutoffs using the optimal population model at N = 291."

getPowerFit(
    h1model_out,
    nullObject = optmodel_out,
    alpha = ALPHA,
    nVal = 291,
    usedFit = FITS
)
```

```{r}
#| label: tbl-fit-cutoff-opt
#| tbl-cap: "Fit index cutoff values using the optimal population model at N = 291."

getCutoff(
    optmodel_out,
    alpha = ALPHA,
    nVal = 291,
    usedFit = FITS
)
```
