---
title: "Simulation — Ideal Conditions"
execute:
  echo: false
  warning: false
  message: false
---

## Setup

```{r}
#| label: setup
#| include: false

library(here)
source(here("Scripts", "DataAppendixScripts", "models.R"))
source(here("Scripts", "DataAppendixScripts", "helpers.R"))
source(here("Scripts", "DataAppendixScripts", "cache_utils.R"))
source(here("Scripts", "DataAppendixScripts", "cache_loader.R"))

cache_load_existing()
```

## Main misspecification simulation

Simulate data from the ideal population model and fit the analysis CFA
model across the sample-size grid in SEQ. This yields the reference
sampling distribution of fit indices under the correctly specified
condition.

```{r}
#| label: sim-ideal-popmodel
#| cache: true

popmodel_ideal <- cache_or_run("popmodel_ideal", {
  sim(
    NULL,
    n = SEQ1,              # See the `SEQ1` object defined in helpers.R for the sample size sequence used in this scenario.
    analyzemodel,
    generate = popmodel,
    std.lv = TRUE,
    lavaanfun = "cfa",
    seed = SEED,
    multicore = TRUE
)
})
```

Simulate data from the alternative population model and fit the same
analysis CFA model across the same sample-size grid. This creates the
sampling distribution of fit indices under misspecification, which is
later contrasted with the reference condition to evaluate power.

```{r}
#| label: sim-ideal-h1model
#| cache: true

h1model_ideal <- cache_or_run("h1model_ideal", {
  sim(
    NULL,
    n = SEQ1,              # See the `SEQ1` object defined in helpers.R for the sample size sequence used in this scenario.
    analyzemodel,
    generate = h1model,
    std.lv = TRUE,
    lavaanfun = "cfa",
    seed = SEED,
    multicore = TRUE
)
})
```

## Main misspecification outputs

### Power curves for fit indices

`plotPowerFit()` displays how the probability of rejecting the
misspecified condition changes as sample size increases, using the
logistic-smoothed relationship between $N$ and the reject/retain
decision for each index [@simsem2021]. The curves cross the target
power level at different sample sizes, reflecting each index's
sensitivity to the imposed cross-loading misspecification.

```{r}
#| label: fig-power-curve-ideal
#| fig-cap: "Power curves for fit indices under ideal conditions."
#| fig-width: 8
#| fig-height: 5

plotPowerFit(
    h1model_ideal,
    popmodel_ideal,
    alpha = ALPHA,
    usedFit = FITS
)
```

The unsmoothed version below overlays the raw simulated fit-index
values from both conditions across the sample-size sequence. This
display is useful for inspecting how the two sampling distributions
separate as $N$ grows, before that separation is summarized into
logistic power curves [@simsem2021; @mcneish2021].

```{r}
#| label: fig-fit-curve-ideal
#| fig-cap: "Power curves for fit indices under ideal conditions without logistic smoothing."
#| fig-width: 8
#| fig-height: 5

plotPowerFit(
    h1model_ideal,
    popmodel_ideal,
    alpha = ALPHA,
    usedFit = FITS,
    logistic = FALSE
)
```

### Power and cutoff values for fit indices

`getPowerFit()` evaluates how often the alternative model is rejected
at a given $N$, whereas `getCutoff()` derives the index-specific
decision boundaries from the null-model sampling distribution at the
same $\alpha$ level. Reading power and cutoffs together is more
informative than either alone: the cutoff defines how stringent each
rejection rule is, and the power shows how often that rule succeeds
[@jobst2023; @wang2023].

```{r}
#| label: tbl-power-cutoff-ideal
#| tbl-cap: "Power and fit index cutoff values under ideal conditions at N = 231."

# Estimate power for rejecting the misspecified model (h1model_ideal)
# using fit-index cutoffs implied by the null model at alpha = ALPHA and N = 231.
power_ideal <- getPowerFit(
    h1model_ideal,
    nullObject = popmodel_ideal,
    alpha = ALPHA,
    nVal = 231,
    usedFit = FITS
)

# Derive empirical cutoff values for the selected fit indices
# from the null-model sampling distribution under the same conditions.
cutoff_ideal <- getCutoff(
    popmodel_ideal,
    alpha = ALPHA,
    nVal = 231,
    usedFit = FITS
)

# Combine power and cutoff values into a compact summary table.
power_cutoff_ideal <- rbind(
    "Power" = as.numeric(power_ideal),
    "Cutoff" = as.numeric(cutoff_ideal[1, ])
)

colnames(power_cutoff_ideal) <- toupper(names(power_ideal)) # Standardize column names for presentation.
knitr::kable(power_cutoff_ideal, digits = 3, align = "c") # Print the table.
```

### Traditional cutoff values for fit indices

The same analysis is repeated using fixed rule-of-thumb thresholds in
place of empirically derived cutoffs. This serves as a direct
illustration of how the choice of decision criterion — not $N$ alone —
affects power conclusions in simulation-based CFA evaluation
[@mcneish2021; @kyriazos2018].

```{r}
#| label: fig-fit-curve-traditional
#| fig-cap: "Power curves using traditional fit index cutoffs under ideal conditions."
#| fig-width: 8
#| fig-height: 5

# Plot fit-index results from the misspecified model across sample sizes.
# The cutoff object applies traditional rule-of-thumb thresholds
# (e.g., RMSEA, SRMR, CFI, and TLI cutoffs) for rejection decisions.
# logistic = FALSE requests the scatterplot-style display of simulated values
# and the cutoff line, instead of a logistic-smoothed power curve

plotPowerFit(
  h1model_ideal,   # Simulated results under the misspecified model
  popmodel_ideal,  # Reference model object used for comparison
  cutoff = RULE_OF_THUMB, # Rule-of-thumb cutoff values for fit indices
  alpha = ALPHA, # Significance level for determining power
  usedFit = FITS, # RMSEA, SRMR, CFI, and TLI fit indices to display
  logistic = FALSE # Display raw simulated points and cutoff lines instead of logistic curves
)
```

The table below summarizes detection rates at $N = 231$ under the same
traditional thresholds, providing a direct comparison with the
empirical-cutoff results from the previous block.

```{r}
#| label: tbl-power-cutoff-traditional
#| tbl-cap: "Power for traditional fit index cutoffs under ideal conditions at N = 231."

power_trad <- getPowerFit(
    h1model_ideal,
    RULE_OF_THUMB,
    alpha = ALPHA,
    nVal = 231,
    usedFit = FITS
)

# Create horizontal table: Fit Index as column names, Power as single row
power_df <- data.frame(
    "CFI" = round(as.numeric(power_trad["cfi"]), 3),
    "TLI" = round(as.numeric(power_trad["tli"]), 3),
    "RMSEA" = round(as.numeric(power_trad["rmsea"]), 3),
    "SRMR" = power_trad["srmr"],  # Keep as string
    row.names = "Power"
)

knitr::kable(power_df, digits = 3, align = "c", col.names = toupper(names(power_df)))
```

### Target parameter estimates

Monte Carlo summaries in `simsem` go beyond detection rates: relative
bias, relative SE bias, and CI coverage jointly indicate whether each
estimator is well-calibrated in finite samples. A parameter may show
high detection power while still being recovered with noticeable bias
or poorly estimated uncertainty, so all four indicators should be
read together rather than treating power alone as the criterion of
interest [@wang2021; @simsem2021].

```{r}
#| label: tbl-param-estimate-ideal
#| tbl-cap: "Summary of parameter estimates under ideal conditions."

# Compute detailed parameter-recovery results for the fitted model.
param_ideal <- summaryParam(
  h1model_ideal,
  alpha = ALPHA,
  digits = 3,
  detail = TRUE
)

# Keep only the diagnostics used in the discussion.
param_subset <- param_ideal[, c("Rel Bias", "Rel SE Bias", "Coverage", "Power (Not equal 0)")]

# Move parameter labels to an explicit first column.
param_subset <- cbind(
  Parameter = rownames(param_subset),
  param_subset,
  row.names = NULL
)

# Build an output-specific display version of the parameter labels.
if (knitr::is_latex_output()) {
  param_subset$Parameter <- gsub(
    "~",
    "\\\\textasciitilde{}",
    param_subset$Parameter,
    fixed = TRUE
  )
  tbl_format <- "latex"
} else if (knitr::is_html_output()) {
  tbl_format <- "html"
} else {
  tbl_format <- "pipe"
}

# Print the table.
knitr::kable(
  param_subset,
  format = tbl_format,
  digits = 3,
  align = c("l", "c", "c", "c", "c"),
  escape = FALSE,
  row.names = FALSE,
  caption = "Parameter estimates under ideal conditions" # REMEMBER ERROR RENDER
)
```

`getPower()` returns parameter-specific power estimates across the
simulated values of $N$, and `findPower()` identifies the first $N$
in the grid that exceeds the target level. The resulting values should
be read as practical planning thresholds: the most demanding parameters
determine the minimum adequate sample size, since planning around the
easiest effects to detect would leave the hardest ones systematically
underpowered [@simsem2021; @moshagen2024].

```{r}
#| label: tbl-power-n-ideal
#| tbl-cap: "Sample size needed for 80% power under ideal conditions."

# Obtain parameter-specific power estimates across sample sizes.
pow_h1 <- getPower(h1model_ideal, alpha = ALPHA)

# Find the minimum N in the simulated grid that reaches the target power of 0.80.
n_h1 <- findPower(pow_h1, "N", 0.80)

# Convert to a data frame for reporting.
n_df <- data.frame(
  Parameter = names(n_h1),
  N = as.integer(n_h1),
  stringsAsFactors = FALSE
)

# Remove parameters for which no finite N was returned.
n_df <- subset(n_df, is.finite(N))

# Build an output-specific display version of the parameter labels.
if (knitr::is_latex_output()) {
  n_df$Parameter <- gsub(
    "~",
    "\\\\textasciitilde{}",
    n_df$Parameter,
    fixed = TRUE
  )
  tbl_format <- "latex"
} else if (knitr::is_html_output()) {
  tbl_format <- "html"
} else {
  tbl_format <- "pipe"
}

# Print the table without row numbers.
knitr::kable(
  n_df,
  format = tbl_format,
  col.names = c("Parameter", "N"),
  align = c("l", "c"),
  escape = FALSE,
  row.names = FALSE
)
```

### Recovering the population parameters

Parameter recovery is a complementary perspective on sample size
planning: instead of asking whether a misspecification can be detected
globally, it asks whether individual model parameters can be
accurately estimated under the assumed data-generating conditions
[@wang2021; @feng2023]. Although primary loadings are typically easy
to recover even at modest $N$, latent covariances and residual
covariances may require substantially larger samples, making them
the practical bottleneck for many CFA designs [@moshagen2024; @gagne2006].

The analyses below use the naive and optimal population models as
sensitivity benchmarks. Because these models differ from the main
meta-analytic population in their loading magnitudes and interfactor
correlations, comparing their recovery profiles clarifies how
sensitive the sample size recommendation is to the assumed parameter
values.

#### Naive model

```{r}
#| label: sim-ideal-naivemodel
#| cache: true

naivemodel_ideal <- cache_or_run("naivemodel_ideal", {
  sim(
    NULL,
    n = SEQ1,               # See the `SEQ1` object defined in helpers.R for the sample size sequence used in this scenario.
    analyzemodel,
    generate = naivemodel,
    std.lv = TRUE,
    lavaanfun = "cfa",
    seed = SEED,
    multicore = TRUE
)
})
```

The table below reports parameter-level diagnostics and the minimum $N$
required for each parameter to reach $.80$ power under the naive
population. Only parameters with a finite required $N$ above 60 are
shown, as smaller requirements are of limited practical interest for
planning purposes [@simsem2021; @gagne2006].

```{r}
#| label: tbl-power-n-naive
#| tbl-cap: "Parameter estimates and required sample size for 80% power under naive conditions."

# Compute detailed parameter-recovery results for the fitted model.
param_naive <- summaryParam(
  naivemodel_ideal,
  alpha = ALPHA,
  digits = 3,
  detail = TRUE
)

# Keep only the diagnostics used in the discussion.
param_subset <- as.data.frame(
  param_naive[, c("Rel Bias", "Rel SE Bias", "Coverage", "Power (Not equal 0)")]
)

# Compute parameter-level power across the simulated values of N.
pow_naive <- getPower(naivemodel_ideal, alpha = ALPHA)

# Extract the minimum N that reaches the target power of .80 for each parameter.
n_naive <- findPower(pow_naive, "N", 0.80)

# Add required N to the subset.
param_subset$N <- as.integer(n_naive[rownames(param_subset)])

# Filter to keep only finite N values greater than 60.
filtered_subset <- subset(param_subset, is.finite(N) & N > 60)

# Move parameter labels to an explicit first column.
filtered_subset <- cbind(
  Parameter = rownames(filtered_subset),
  filtered_subset,
  row.names = NULL
)

# Build an output-specific display version of the parameter labels.
if (knitr::is_latex_output()) {
  filtered_subset$Parameter <- gsub(
    "~",
    "\\\\textasciitilde{}",
    filtered_subset$Parameter,
    fixed = TRUE
  )
  tbl_format <- "latex"
} else if (knitr::is_html_output()) {
  tbl_format <- "html"
} else {
  tbl_format <- "pipe"
}

# Print the combined table.
knitr::kable(
  filtered_subset,
  format = tbl_format,
  digits = 3,
  align = c("l", "c", "c", "c", "c", "c"),
  escape = FALSE,
  col.names = c(
    "Parameter",
    "Rel Bias",
    "Rel SE Bias",
    "Coverage",
    "Power (Not equal 0)",
    "Required N"
  )
)
```

#### Optimal model

```{r}
#| label: sim-ideal-optmodel
#| cache: true

optmodel_ideal <- cache_or_run("optmodel_ideal", {
  sim(
    NULL,
    n = SEQ1,               # See the `SEQ1` object defined in helpers.R for the sample size sequence used in this scenario.
    analyzemodel,
    generate = optmodel,
    std.lv = TRUE,
    lavaanfun = "cfa",
    seed = SEED,
    multicore = TRUE
)
})
```

The same parameter-level analysis is applied to the optimal model.
Because this population uses stronger loadings and lower interfactor
correlations, most parameters are expected to reach the target power
at smaller $N$ values than under the naive scenario, and the set of
finite requirements will likely be sparser [@simsem2021; @buchberger2024].

```{r}
#| label: tbl-power-n-opt
#| tbl-cap: "Parameter estimates and required sample size for 80% power under optimal conditions."

# Compute detailed parameter-recovery results for the fitted model.
param_opt <- summaryParam(
  optmodel_ideal,
  alpha = ALPHA,
  digits = 3,
  detail = TRUE
)

# Keep only the diagnostics used in the discussion.
param_subset <- as.data.frame(
  param_opt[, c("Rel Bias", "Rel SE Bias", "Coverage", "Power (Not equal 0)")]
)

# Compute parameter-level power across the simulated values of N.
pow_opt <- getPower(optmodel_ideal, alpha = ALPHA)

# Extract the minimum N that reaches the target power of .80 for each parameter.
n_opt <- findPower(pow_opt, "N", 0.80)

# Add required N to the subset.
param_subset$N <- as.integer(n_opt[rownames(param_subset)])

# Filter to keep only finite N values greater than 60.
filtered_subset <- subset(param_subset, is.finite(N) & N > 60)

# Move parameter labels to an explicit first column.
filtered_subset <- cbind(
  Parameter = rownames(filtered_subset),
  filtered_subset,
  row.names = NULL
)

# Build an output-specific display version of the parameter labels.
if (knitr::is_latex_output()) {
  filtered_subset$Parameter <- gsub(
    "~",
    "\\\\textasciitilde{}",
    filtered_subset$Parameter,
    fixed = TRUE
  )
  tbl_format <- "latex"
} else if (knitr::is_html_output()) {
  tbl_format <- "html"
} else {
  tbl_format <- "pipe"
}

# Print the combined table.
knitr::kable(
  filtered_subset,
  format = tbl_format,
  digits = 3,
  align = c("l", "c", "c", "c", "c", "c"),
  escape = FALSE,
  col.names = c(
    "Parameter",
    "Rel Bias",
    "Rel SE Bias",
    "Coverage",
    "Power (Not equal 0)",
    "Required N"
  )
)
```

### Misspecification and population sensitivity

#### Naive model

The chunks below replicate the main misspecification analysis with
the naive model as the reference population. Replacing `popmodel`
with `naivemodel` tests how sensitive the power conclusions are to
the assumed parameter values: weaker loadings and higher interfactor
correlations reduce the signal available to each fit index, typically
pushing the required $N$ upward [@buchberger2024; @irmer2024].

```{r}
#| label: fig-power-curve-naive
#| fig-cap: "Power curves for fit indices using the naive population model."
#| fig-width: 8
#| fig-height: 5

# Plot fit-index-based power as a function of N.
# With logistic = TRUE (default), simsem smooths the reject/retain decisions
# across sample size using logistic regression, yielding continuous power curves
# for each fit index.
plotPowerFit(
  h1model_ideal,
  naivemodel_ideal,
  alpha = ALPHA,
  usedFit = FITS
)
```

```{r}
#| label: fig-fit-curve-naive
#| fig-cap: "Power curves for fit indices using the naive population model without logistic smoothing."
#| fig-width: 8
#| fig-height: 5

# Plot the raw fit-index distributions over N without logistic smoothing.
# This version is useful to inspect the empirical separation between the null
# and alternative model results before smoothing them into power curves.
plotPowerFit(
  h1model_ideal,
  naivemodel_ideal,
  alpha = ALPHA,
  usedFit = FITS,
  logistic = FALSE
)
```

The table below summarizes power and empirical cutoffs at $N = 330$
under the naive reference population.

```{r}
#| label: tbl-power-cutoff-naive
#| tbl-cap: "Power and fit index cutoff values using the naive population model at N = 330."

# Estimate the power to reject the misspecified model at N = 330 for each fit index.
power_naive <- getPowerFit(
  h1model_ideal,
  nullObject = naivemodel_ideal,
  alpha = ALPHA,
  nVal = 330,
  usedFit = FITS
)

# Derive empirical fit-index cutoffs from the null model at the same alpha level.
cutoff_naive <- getCutoff(
  naivemodel_ideal,
  alpha = ALPHA,
  nVal = 330,
  usedFit = FITS
)

# Combine power and cutoff values into a compact summary table.
power_cutoff_naive <- rbind(
  "Power" = as.numeric(power_naive),
  "Cutoff" = as.numeric(cutoff_naive[1, ])
)
colnames(power_cutoff_naive) <- toupper(names(power_naive))

# Print the final table.
knitr::kable(power_cutoff_naive, digits = 3, align = "c")
```

#### Optimal model

The same sensitivity check is repeated with the optimal model as the
reference population. Stronger loadings and lower interfactor
correlations are expected to shift the power curves leftward, reducing
the $N$ needed for adequate discrimination relative to the naive
scenario [@mosqueirataipe2026; @buchberger2024].

```{r}
#| label: fig-power-curve-opt
#| fig-cap: "Power curves for fit indices using the optimal population model."
#| fig-width: 8
#| fig-height: 5

# Plots power as a function of N for the selected fit indices.
# The null/reference object provides the cutoff distribution.
# With the default logistic = TRUE, the curves are smoothed.

plotPowerFit(
  h1model_ideal,
  optmodel_ideal,
  alpha = ALPHA,
  usedFit = FITS
)
```

```{r}
#| label: fig-fit-curve-opt
#| fig-cap: "Power curves for fit indices using the optimal population model without logistic smoothing."
#| fig-width: 8
#| fig-height: 5

# Repeats the previous plot without logistic smoothing.
# This version emphasizes the raw simulated fit-index patterns across N.

plotPowerFit(
  h1model_ideal,
  optmodel_ideal,
  alpha = ALPHA,
  usedFit = FITS,
  logistic = FALSE
)
```

The table below summarizes power and empirical cutoffs at $N = 291$
under the optimal reference population.

```{r}
#| label: tbl-power-cutoff-opt
#| tbl-cap: "Power and fit index cutoff values using the optimal population model at N = 291."

# Power at the target sample size for each fit index.
power_opt <- getPowerFit(
  h1model_ideal,
  nullObject = optmodel_ideal,
  alpha = ALPHA,
  nVal = 291,
  usedFit = FITS
)

# Index-specific cutoffs derived from the reference model at the same N.
cutoff_opt <- getCutoff(
  optmodel_ideal,
  alpha = ALPHA,
  nVal = 291,
  usedFit = FITS
)

# Combines power and cutoff values into a compact display table.
power_cutoff_opt <- rbind(
  "Power" = as.numeric(power_opt),
  "Cutoff" = as.numeric(cutoff_opt[1, ])
)
colnames(power_cutoff_opt) <- toupper(names(power_opt))

knitr::kable(power_cutoff_opt, digits = 3, align = "c")
```
