---
title: "Simulation — Ideal Conditions"
execute:
  echo: false
  warning: false
  message: false
---

## Setup

```{r}
#| label: setup
#| include: false

# For CFA population models
source(here::here("Scripts", "DataAppendixScripts", "models.R"))

# For DRY (Don't repeat yourself) parameters
source(here::here("Scripts", "DataAppendixScripts", "helpers.R"))

# For simulations caching
source(here::here("Scripts", "DataAppendixScripts", "cache_utils.R"))

# For loading existing simulations cache
source(here::here("Scripts", "DataAppendixScripts", "cache_loader.R"))

# Load existing simulations cache if available (cache_loader.R)
cache_load_existing()
```

## Main misspecification simulation

Simulate data from the ideal population model and fit the analysis CFA
model across the sample-size grid in `SEQ1` (see `helpers.R`). This yields the reference
sampling distribution of fit indices under the correctly specified
condition.

```{r}
#| label: sim-ideal-popmodel

popmodel_ideal <- cache_or_run("popmodel_ideal", {
  sim(
    NULL,                  # Number of replications is defined by SEQ1
    n = SEQ1,              # Sample sizes defined in SEQ1 (see helpers.R)
    analyzemodel,          # Model used for analysis (see models.R)
    generate = popmodel,   # Population model (see models.R)
    std.lv = TRUE,         # Standardize latent variables
    lavaanfun = "cfa",     # Use cfa() function for analysis
    seed = SEED,           # Seed for reproducibility (see helpers.R)
    multicore = TRUE       # Use multiple cores for parallel processing
)
})
```

Simulate data from the alternative population model and fit the same
analysis CFA model across the same sample-size grid. This creates the
sampling distribution of fit indices under misspecification, which is
later contrasted with the reference condition to evaluate power.

```{r}
#| label: sim-ideal-h1model

h1model_ideal <- cache_or_run("h1model_ideal", {
  sim(
    NULL,               # Number of replications is defined by SEQ1 
    n = SEQ1,           # Sample sizes defined in SEQ1
    analyzemodel,       # Model used for analysis
    generate = h1model, # Population model (plus 3 cross-loadings)
    std.lv = TRUE,      # Standardize latent variables
    lavaanfun = "cfa",  # Use cfa() function for analysis
    seed = SEED,        # Seed for reproducibility
    multicore = TRUE    # Use multiple cores for parallel processing
)
})
```

## Main misspecification outputs

### Power curves for fit indices

`plotPowerFit()` displays how the probability of rejecting the
misspecified condition changes as sample size increases, using the
logistic-smoothed relationship between $N$ and the reject/retain
decision for each index [@simsem2021]. The curves cross the target
power level at different sample sizes, reflecting each index's
sensitivity to the imposed cross-loading misspecification.

```{r}
#| label: fig-power-curve-ideal
#| fig-cap: "Power curves for fit indices under ideal conditions."
#| fig-width: 8
#| fig-height: 5

plotPowerFit(
    h1model_ideal,   # Data simulated under H1 (with cross-loadings)
    popmodel_ideal,  # Data simulated under H0 (no cross-loadings)
    alpha = ALPHA,   # Significance level for determining power (see helpers.R)
    usedFit = FITS   # RMSEA, SRMR, CFI, and TLI fit indices to display (see helpers.R)
)
```

The unsmoothed version below overlays the raw simulated fit-index
values from both conditions across the sample-size sequence. This
display is useful for inspecting how the two sampling distributions
separate as $N$ grows, before that separation is summarized into
logistic power curves [@simsem2021; @mcneish2021].

```{r}
#| label: fig-fit-curve-ideal
#| fig-cap: "Power curves for fit indices under ideal conditions without logistic smoothing."
#| fig-width: 8
#| fig-height: 5

plotPowerFit(
    h1model_ideal,    # Data simulated under H1 (with cross-loadings)
    popmodel_ideal,   # Data simulated under H0 (no cross-loadings)
    alpha = ALPHA,    # Significance level for determining power (see helpers.R)
    usedFit = FITS,   # RMSEA, SRMR, CFI, and TLI fit indices to display (see helpers.R)
    logistic = FALSE  # Display raw simulated points and cutoff lines instead of logistic curves
)
```

### Power and cutoff values for fit indices

`getPowerFit()` evaluates how often the alternative model is rejected
at a given $N$, whereas `getCutoff()` derives the index-specific
decision boundaries from the null-model sampling distribution at the
same $\alpha$ level. Reading power and cutoffs together is more
informative than either alone: the cutoff defines how stringent each
rejection rule is, and the power shows how often that rule succeeds
[@jobst2023; @wang2023].

```{r}
#| label: tbl-power-cutoff-ideal
#| tbl-cap: "Power and fit index cutoff values under ideal conditions at N = 231."

# Estimate power for rejecting the misspecified model (h1model_ideal)
# using fit-index cutoffs implied by the null model (popmodel_ideal) at alpha = ALPHA and N = 231.
power_ideal <- getPowerFit(
    h1model_ideal,                # Data simulated under H1 (with cross-loadings)
    nullObject = popmodel_ideal,  # Data simulated under H0 (no cross-loadings)
    alpha = ALPHA,                # Significance level for determining power (see helpers.R)
    nVal = 231,                   # Sample size for analysis
    usedFit = FITS                # RMSEA, SRMR, CFI, and TLI fit indices to display (see helpers.R)
)

# Derive empirical cutoff values for the selected fit indices
# from the null-model sampling distribution under the same conditions.
cutoff_ideal <- getCutoff(
    popmodel_ideal,               # Data simulated under H0 (no cross-loadings)
    alpha = ALPHA,                # Significance level for determining power (see helpers.R)
    nVal = 231,                   # Sample size for analysis
    usedFit = FITS                # RMSEA, SRMR, CFI, and TLI fit indices to display (see helpers.R)
)

# Combine power and cutoff values into a compact summary table.
power_cutoff_ideal <- rbind(
    "Power" = as.numeric(power_ideal),
    "Cutoff" = as.numeric(cutoff_ideal[1, ])
)

# Standardize column names for presentation.
colnames(power_cutoff_ideal) <- toupper(names(power_ideal)) 

# Print the table.
knitr::kable(power_cutoff_ideal, digits = 3, align = "c") 
```

### Traditional cutoff values for fit indices

The same analysis is repeated using fixed rule-of-thumb thresholds in
place of empirically derived cutoffs. This serves as a direct
illustration of how the choice of decision criterion — not $N$ alone —
affects power conclusions in simulation-based CFA evaluation
[@mcneish2021; @kyriazos2018].

```{r}
#| label: fig-fit-curve-traditional
#| fig-cap: "Power curves using traditional fit index cutoffs under ideal conditions."
#| fig-width: 8
#| fig-height: 5

# Plot fit-index results from the misspecified model across sample sizes.
# The cutoff object applies traditional rule-of-thumb thresholds
# (e.g., RMSEA, SRMR, CFI, and TLI cutoffs) for rejection decisions.
# logistic = FALSE requests the scatterplot-style display of simulated values
# and the cutoff line, instead of a logistic-smoothed power curve

plotPowerFit(
  h1model_ideal,          # Data simulated under H1 (with cross-loadings)
  popmodel_ideal,         # Data simulated under H0 (no cross-loadings)
  cutoff = RULE_OF_THUMB, # Rule-of-thumb cutoff values for fit indices (see helpers.R)
  alpha = ALPHA,          # Significance level for determining power (see helpers.R)
  usedFit = FITS,         # RMSEA, SRMR, CFI, and TLI fit indices to display (see helpers.R)
  logistic = FALSE        # Display raw simulated points and cutoff lines instead of logistic curves
)
```

The table below summarizes detection rates at $N = 231$ under the same
traditional thresholds, providing a direct comparison with the
empirical-cutoff results from the previous block.

```{r}
#| label: tbl-power-cutoff-traditional-ideal
#| tbl-cap: "Power for traditional fit index cutoffs under ideal conditions at N = 231."

# Estimate power for rejecting the misspecified model (h1model_ideal)
# using traditional rule-of-thumb cutoffs (RULE_OF_THUMB) at alpha = ALPHA and N = 231.
power_trad <- getPowerFit(
    h1model_ideal,    # Data simulated under H1 (with cross-loadings)
    RULE_OF_THUMB,    # Rule-of-thumb cutoff values for fit indices (see helpers.R)
    alpha = ALPHA,    # Significance level for determining power (see helpers.R)
    nVal = 231,       # Sample size for analysis
    usedFit = FITS    # RMSEA, SRMR, CFI, and TLI fit indices to display (see helpers.R)
)

# Create horizontal table: Fit Index as column names, Power as single row
power_df <- data.frame(
    "CFI" = round(as.numeric(power_trad["cfi"]), 3),
    "TLI" = round(as.numeric(power_trad["tli"]), 3),
    "RMSEA" = round(as.numeric(power_trad["rmsea"]), 3),
    "SRMR" = power_trad["srmr"],
    row.names = "Power"
)

# Print the table.
knitr::kable(power_df, digits = 3, align = "c", col.names = toupper(names(power_df)))
```

### Target parameter estimates

Monte Carlo summaries in `simsem` go beyond detection rates: relative
bias, relative SE bias, and CI coverage jointly indicate whether each
estimator is well-calibrated in finite samples. A parameter may show
high detection power while still being recovered with noticeable bias
or poorly estimated uncertainty, so all four indicators should be
read together rather than treating power alone as the criterion of
interest [@wang2021; @simsem2021].

```{r}
#| label: tbl-param-estimate-ideal
#| tbl-cap: "Summary of parameter estimates under ideal conditions."

# Compute detailed parameter-recovery results for the fitted model.
param_ideal <- summaryParam(
  h1model_ideal,   # Data simulated under H1 (with cross-loadings)
  alpha = ALPHA,   # Significance level for determining power (see helpers.R)
  digits = 3,      # Number of decimal places for output
  detail = TRUE    # Display detailed parameter-recovery results
)

# Keep only the diagnostics used in the discussion.
param_subset <- param_ideal[, c("Rel Bias", "Rel SE Bias", "Coverage", "Power (Not equal 0)")]

# Move parameter labels to an explicit first column.
param_subset <- cbind(
  Parameter = rownames(param_subset),
  param_subset,
  row.names = NULL
)

# Escape ~~ for Pandoc (prevents strikethrough in HTML); matches tbl-pop-summary
param_subset$Parameter <- gsub("~~", "\\~\\~", param_subset$Parameter, fixed = TRUE)

# Print the table.
knitr::kable(
  param_subset,
  digits = 3,
  align = c("l", "c", "c", "c", "c"),
  booktabs = TRUE,
  escape = FALSE,
  row.names = FALSE
)
```

`getPower()` returns parameter-specific power estimates across the
simulated values of $N$, and `findPower()` identifies the first $N$
in the grid that exceeds the target level. The resulting values should
be read as practical planning thresholds: the most demanding parameters
determine the minimum adequate sample size, since planning around the
easiest effects to detect would leave the hardest ones systematically
underpowered [@simsem2021; @moshagen2024].

```{r}
#| label: tbl-power-n-ideal
#| warning: false
#| tbl-cap: "Sample size needed for 80% power under ideal conditions."

# Obtain parameter-specific power estimates across sample sizes.
pow_h1 <- getPower(h1model_ideal, alpha = ALPHA)

# Find the minimum N in the simulated grid that reaches the target power of 0.80.
n_h1 <- findPower(pow_h1, "N", 0.80)

# Convert to a data frame for reporting.
n_df <- data.frame(
  Parameter = names(n_h1),
  N = suppressWarnings(as.integer(n_h1)),
  stringsAsFactors = FALSE
)

# Remove parameters for which no finite N was returned.
n_df <- subset(n_df, is.finite(N))

# Escape ~~ for Pandoc (prevents strikethrough in HTML); matches tbl-pop-summary
n_df$Parameter <- gsub("~~", "\\~\\~", n_df$Parameter, fixed = TRUE)

# Print the table without row numbers.
knitr::kable(
  n_df,
  col.names = c("Parameter", "N"),
  align = c("l", "c"),
  booktabs = TRUE,
  escape = FALSE,
  row.names = FALSE
)
```

### Recovering the population parameters

Parameter recovery is a complementary perspective on sample size
planning: instead of asking whether a misspecification can be detected
globally, it asks whether individual model parameters can be
accurately estimated under the assumed data-generating conditions
[@wang2021; @feng2023]. Although primary loadings are typically easy
to recover even at modest $N$, latent covariances and residual
covariances may require substantially larger samples, making them
the practical bottleneck for many CFA designs [@moshagen2024; @gagne2006].

The analyses below use the naive and optimal population models as
sensitivity benchmarks. Because these models differ from the main
meta-analytic population in their loading magnitudes and interfactor
correlations, comparing their recovery profiles clarifies how
sensitive the sample size recommendation is to the assumed parameter
values.

#### Naive model

```{r}
#| label: sim-ideal-naivemodel

naivemodel_ideal <- cache_or_run("naivemodel_ideal", {
  sim(
    NULL,                   # Number of replications is defined by SEQ1
    n = SEQ1,               # Sample sizes defined in SEQ1 (see helpers.R)
    analyzemodel,           # Model used for analysis (see models.R)
    generate = naivemodel,  # Naive population model (see models.R)
    std.lv = TRUE,          # Standardized latent variables
    lavaanfun = "cfa",      # Use CFA function from lavaan
    seed = SEED,            # Seed for reproducibility (see helpers.R)
    multicore = TRUE        # Use multicore processing
)
})
```

The table below reports parameter-level diagnostics and the minimum $N$
required for each parameter to reach $.80$ power under the naive
population. Only parameters with a finite required $N$ above 60 are
shown, as smaller requirements are of limited practical interest for
planning purposes [@simsem2021; @gagne2006].

```{r}
#| label: tbl-power-n-naive
#| warning: false
#| message: false
#| tbl-cap: "Parameter estimates and required sample size for 80% power under naive conditions."

# Compute detailed parameter-recovery results for the fitted model.
param_naive <- summaryParam(
  naivemodel_ideal,  # Simulation results from the naive population model (low loadings)
  alpha = ALPHA,     # Significance level for determining power (see helpers.R)
  digits = 3,        # Number of digits to round the results to
  detail = TRUE      # Return detailed parameter-level results
)

# Keep only the diagnostics used in the discussion.
param_subset <- as.data.frame(
  param_naive[, c("Rel Bias", "Rel SE Bias", "Coverage", "Power (Not equal 0)")]
)

# Compute parameter-level power across the simulated values of N.
pow_naive <- getPower(naivemodel_ideal, alpha = ALPHA)

# Extract the minimum N that reaches the target power of .80 for each parameter.
n_naive <- findPower(pow_naive, "N", 0.80)

# Add required N to the subset.
param_subset$N <- suppressWarnings(as.integer(n_naive[rownames(param_subset)]))

# Filter to keep only finite N values greater than 60.
filtered_subset <- subset(param_subset, is.finite(N) & N > 60)

# Move parameter labels to an explicit first column.
filtered_subset <- cbind(
  Parameter = rownames(filtered_subset),
  filtered_subset,
  row.names = NULL
)

# Escape ~~ for Pandoc (prevents strikethrough in HTML); matches tbl-pop-summary
filtered_subset$Parameter <- gsub("~~", "\\~\\~", filtered_subset$Parameter, fixed = TRUE)

# Print the combined table.
knitr::kable(
  filtered_subset,
  digits = 3,
  align = c("l", "c", "c", "c", "c", "c"),
  booktabs = TRUE,
  escape = FALSE,
  col.names = c(
    "Parameter",
    "Rel Bias",
    "Rel SE Bias",
    "Coverage",
    "Power (Not equal 0)",
    "Required N"
  )
)
```

#### Optimal model

```{r}
#| label: sim-ideal-optmodel

optmodel_ideal <- cache_or_run("optmodel_ideal", {
  sim(
    NULL,                   # Number of replications is defined by SEQ1
    n = SEQ1,               # Sample sizes defined in SEQ1 (see helpers.R)
    analyzemodel,           # Model used for analysis (see models.R)
    generate = optmodel,    # Optimal population model (see models.R)
    std.lv = TRUE,          # Standardized latent variables
    lavaanfun = "cfa",      # Use CFA function from lavaan
    seed = SEED,            # Seed for reproducibility (see helpers.R)
    multicore = TRUE        # Use multicore processing
)
})
```

The same parameter-level analysis is applied to the optimal model.
Because this population uses stronger loadings ($\lambda \approx 0.80$)
and lower interfactor correlations ($\phi = 0.08$), two contrasting phenomena
emerge. On the one hand, primary factor loadings and residual variances reach
the target power of $0.80$ at minimal sample sizes ($N \le 51$) with negligible bias.
On the other hand, the very small interfactor correlations ($\phi = 0.08$) create
an acute bottleneck: detecting such weak associations with $80\%$ power requires
$N > 1,200$, far exceeding the simulated range ($N \in [51, 350]$), which causes
`findPower` to return `NA` for all factor correlations. Consequently, among parameters
with finite sample size requirements above 60, only the residual covariance
$Q3 \sim\sim Q4$ appears in the table below, requiring $N = 139$ [@simsem2021; @buchberger2024].

```{r}
#| label: tbl-power-n-opt
#| warning: false
#| message: false
#| tbl-cap: "Parameter estimates and required sample size for 80% power under optimal conditions."

# Compute detailed parameter-recovery results for the fitted model.
param_opt <- summaryParam(
  optmodel_ideal,   # Simulation results from the optimal population model (stronger factor loadings)
  alpha = ALPHA,    # Significance level for determining power (see helpers.R)
  digits = 3,       # Number of digits to round the results to
  detail = TRUE     # Return detailed parameter-level results
)

# Keep only the diagnostics used in the discussion.
param_subset <- as.data.frame(
  param_opt[, c("Rel Bias", "Rel SE Bias", "Coverage", "Power (Not equal 0)")]
)

# Compute parameter-level power across the simulated values of N.
pow_opt <- getPower(optmodel_ideal, alpha = ALPHA)

# Extract the minimum N that reaches the target power of .80 for each parameter.
n_opt <- findPower(pow_opt, "N", 0.80)

# Add required N to the subset.
param_subset$N <- suppressWarnings(as.integer(n_opt[rownames(param_subset)]))

# Filter to keep only finite N values greater than 60.
filtered_subset <- subset(param_subset, is.finite(N) & N > 60)

# Move parameter labels to an explicit first column.
filtered_subset <- cbind(
  Parameter = rownames(filtered_subset),
  filtered_subset,
  row.names = NULL
)

# Escape ~~ for Pandoc (prevents strikethrough in HTML); matches tbl-pop-summary
filtered_subset$Parameter <- gsub("~~", "\\~\\~", filtered_subset$Parameter, fixed = TRUE)

# Print the combined table.
knitr::kable(
  filtered_subset,
  digits = 3,
  align = c("l", "c", "c", "c", "c", "c"),
  booktabs = TRUE,
  escape = FALSE,
  col.names = c(
    "Parameter",
    "Rel Bias",
    "Rel SE Bias",
    "Coverage",
    "Power (Not equal 0)",
    "Required N"
  )
)
```

## References {.unnumbered}

::: {#refs}
:::

