---
title: "Simulation — Ideal Conditions"
execute:
  echo: false
---

## Setup

```{r}
#| label: setup
#| include: false

library(here)
source(here("Scripts", "DataAppendixScripts", "helpers.R"))
source(here("Scripts", "DataAppendixScripts", "models.R"))
```


## Fixed-N simulation (continuous data, ML)

Under ideal conditions: continuous multivariate normal data, ML estimator.
This should converge with the analytical results from semPower/semTools.

```{r}
#| label: sim-ideal-fixed
#| cache: true

sim_ideal_fixed <- run_sim(
  nRep     = N_REP,
  n        = 500,
  model    = analyzeModel,
  generate = popModel
)

summary(sim_ideal_fixed)
```

## Varying-N simulation — Power curve

```{r}
#| label: sim-ideal-vary
#| cache: true

sim_ideal_vary <- run_sim(
  nRep     = NULL,
  n        = N_SEQ,
  model    = analyzeModel,
  generate = popModel
)

summary(sim_ideal_vary)
```

## Power for parameters

```{r}
#| label: sim-ideal-power

# Extract power for each parameter across N values
ideal_power <- getPower(sim_ideal_vary, alpha = ALPHA)

# Find the minimum N for 80% power across all parameters
findPower(ideal_power, "N", POWER)
```

## Power to detect good fit (fit index cutoffs)

```{r}
#| label: sim-ideal-fit-power

# Power that fit indices fall within rule-of-thumb cutoffs at selected N values
cat("--- Power at N = 300 ---\n")
getPowerFit(sim_ideal_vary, cutoff = RULE_OF_THUMB, nVal = 300)

cat("\n--- Power at N = 500 ---\n")
getPowerFit(sim_ideal_vary, cutoff = RULE_OF_THUMB, nVal = 500)

cat("\n--- Power at N = 1000 ---\n")
getPowerFit(sim_ideal_vary, cutoff = RULE_OF_THUMB, nVal = 1000)
```

## Fit index sampling distributions

```{r}
#| label: fig-sim-ideal-fit
#| fig-cap: "Sampling distributions of fit indices across varying N (ideal conditions)"
#| fig-width: 10
#| fig-height: 8

plotCutoff(sim_ideal_vary, alpha = ALPHA)
```

```{r}
#| label: fig-sim-ideal-power-fit
#| fig-cap: "Power to achieve acceptable fit across varying N (ideal conditions)"
#| fig-width: 10
#| fig-height: 6

plotPowerFit(sim_ideal_vary, cutoff = RULE_OF_THUMB, logistic = FALSE)
```

## Parameter-level summary

```{r}
#| label: tbl-sim-param
#| tbl-cap: "Parameter estimates summary — Ideal conditions, fixed N = 500"

param_summary <- summaryParam(sim_ideal_fixed, detail = TRUE, digits = 3)
kable(param_summary)
```

## Convergence: Analytical vs. Simulation

```{r}
#| label: tbl-convergence
#| tbl-cap: "Convergence check — Analytical vs. simulation N requirements"

# Note: fill in analytical_n after running 03_analytical_power.qmd
# For now, use placeholder that will be replaced by the embed output
convergence_table <- data.frame(
  Approach = c("semPower a priori (RMSEA ≥ .05)",
               "semTools a priori (RMSEA ≥ .05)",
               "simsem (parameter power, max N)",
               "simsem (fit power at SRMR < .06)"),
  `N for 80% Power` = c(
    "See @sec-apriori",
    "See @sec-apriori",
    "From findPower() above",
    "From getPowerFit() above"
  ),
  Notes = c(
    "Model-free, RMSEA-based",
    "Same underlying math",
    "Continuous/ML, population model",
    "Continuous/ML, population model"
  ),
  check.names = FALSE
)

kable(convergence_table, align = "lll")
```
