---
title: "Simulation — Ideal Conditions"
execute:
  echo: false
---

## Setup

```{r}
#| label: setup
#| include: false

library(here)
source(here("Scripts", "DataAppendixScripts", "helpers.R"))
source(here("Scripts", "DataAppendixScripts", "models.R"))
```

## Main misspecification simulation

Coul be create an trivial (or harder) artificial misspecification

```{r}
#| label: sim-ideal-popmodel
#| cache: true

popmodel_out <- sim(
    NULL,
    n = 51:350,
    analyzemodel,
    generate = popmodel,
    std.lv = TRUE,
    lavaanfun = "cfa",
    seed = SEED,
    multicore = TRUE
)
```


```{r}
#| label: sim-ideal-h1model
#| cache: true

h1model_out <- sim(
    NULL,
    n = 51:350,
    analyzemodel,
    generate = h1model,
    std.lv = TRUE,
    lavaanfun = "cfa",
    seed = SEED,
    multicore = TRUE
)
```

## Main misspecification outputs

### Power curves for fit indices

```{r}
#| label: power-curve-ideal

plotPowerFit(
    h1model_out,
    popmodel_out,
    alpha = ALPHA,
    usedFit = FITS
)
```


```{r}
#| label: fit-curve-ideal

plotPowerFit(
    h1model_out,
    popmodel_out,
    alpha = ALPHA,
    usedFit = FITS,
    logistic = FALSE
)
```

### Power cutoff values for fit indices

You extract the cutoff values for fit indices at a given alpha level and sample size to determine the threshold for acceptable fit under the null model. This allows you to see how stringent the fit criteria are at different sample sizes.

```{r}
#| label: power-cutoff-ideal

getPowerFit(
    h1model_out,
    nullObject = popmodel_out,
    alpha = ALPHA,
    nVal = 178,
    usedFit = FITS
)
```

```{r}
#| label: fit-cutoff-ideal

getCutoff(
    popmodel_out,
    alpha = ALPHA,
    nVal = 178,
    usedFit = FITS
)
```

### Tradidional cutoff values for fit indices


```{r}
#| label: fit-curve-traditional

plotPowerFit(
    h1model_out,
    popmodel_out,
    cutoff = RULE_OF_THUMB,
    alpha = ALPHA,
    usedFit = FITS,
    logistic = FALSE
)
```


```{r}
#| label: power-cutoff-traditional

getPowerFit(
    h1model_out,
    RULE_OF_THUMB,
    alpha = ALPHA,
    nVal = 178,
    usedFit = FITS
)
```

```{r}
summaryFit(h1model_out, alpha = ALPHA, usedFit = FITS)
```

### Target parameter estimates


```{r}
summaryParam(h1model_out, alpha = ALPHA, digits = 3, detail = TRUE)
```

```{r}
pow_h1 <- getPower(h1model_out, alpha = ALPHA)
findPower(pow_h1, "N", 0.80)
```

### Recovering the population parameters

Recovery of population parameters is not best strategy in CFA power analyses. The idea is to determine the sample size needed to accurately estimate the population parameters (e.g., factor loadings <> 0) with a certain level of precision. This can be done by simulating data under different models and evaluating how well the analysis model recovers the population parameters. 

But this strategy its not pratical because uselly we don't need an large sample size to recover the loadings > 0, but rather to detect misspecifications (e.g., cross-loadings) that are not present in the population model. Therefore, we can use the power to detect misspecifications as a more relevant criterion for determining the required sample size.

Naive model: tau-equivalent loadings (uniform λ per domain) with same reliability as the population model. This is a common but often unrealistic assumption in CFA power analyses.

```{r}
#| label: sim-ideal-naivemodel
#| cache: true

naivemodel_out <- sim(
    NULL,
    n = 51:350,
    analyzemodel,
    generate = naivemodel,
    std.lv = TRUE,
    lavaanfun = "cfa",
    seed = SEED,
    multicore = TRUE
)
```

Maybe the better strategy is recovering N from the lowest population loadings.

```{r}
pow_naive <- getPower(naivemodel_out, alpha = ALPHA)
findPower(pow_naive, "N", 0.80)
```

Optimal model: tau-equivalent loadings (uniform λ per domain) with same reliability as the population model, but with higher loadings to match the population reliability. This represents an idealized scenario where the analysis model is correctly specified in terms of reliability, but still assumes tau-equivalence.

```{r}
#| label: sim-ideal-optmodel
#| cache: true

optmodel_out <- sim(
    NULL,
    n = 51:350,
    analyzemodel,
    generate = optmodel,
    std.lv = TRUE,
    lavaanfun = "cfa",
    seed = SEED,
    multicore = TRUE
)
```

Maybe the better strategy is recovering N from the lowest population loadings.

```{r}
pow_opt <- getPower(optmodel_out, alpha = ALPHA)
findPower(pow_opt, "N", 0.80)
```

### Misspecification simulation with different population parameters

#### Naive model

- Power curves for fit indices

```{r}
#| label: power-curve-naive

plotPowerFit(
    h1model_out,
    naivemodel_out,
    alpha = ALPHA,
    usedFit = FITS
)
```


```{r}
#| label: fit-curve-naive

plotPowerFit(
    h1model_out,
    naivemodel_out,
    alpha = ALPHA,
    usedFit = FITS,
    logistic = FALSE
)
```

- Power cutoff values for fit indices

```{r}
#| label: power-cutoff-naive

getPowerFit(
    h1model_out,
    nullObject = naivemodel_out,
    alpha = ALPHA,
    nVal = 245,
    usedFit = FITS
)
```

```{r}
#| label: fit-cutoff-naive

getCutoff(
    naivemodel_out,
    alpha = ALPHA,
    nVal = 245,
    usedFit = FITS
)
```

#### Optimal model

- Power curves for fit indices

```{r}
#| label: power-curve-opt

plotPowerFit(
    h1model_out,
    optmodel_out,
    alpha = ALPHA,
    usedFit = FITS
)
```


```{r}
#| label: fit-curve-opt

plotPowerFit(
    h1model_out,
    optmodel_out,
    alpha = ALPHA,
    usedFit = FITS,
    logistic = FALSE
)
```

- Power cutoff values for fit indices

```{r}
#| label: power-cutoff-opt

getPowerFit(
    h1model_out,
    nullObject = optmodel_out,
    alpha = ALPHA,
    nVal = 217,
    usedFit = FITS
)
```

```{r}
#| label: fit-cutoff-opt

getCutoff(
    optmodel_out,
    alpha = ALPHA,
    nVal = 217,
    usedFit = FITS
)
```
