---
title: "Simulation — Ideal Conditions"
execute:
  echo: false
---

## Setup

```{r}
#| label: setup
#| include: false

library(here)
source(here("Scripts", "DataAppendixScripts", "helpers.R"))
source(here("Scripts", "DataAppendixScripts", "models.R"))
source(here("Scripts", "DataAppendixScripts", "cache_utils.R"))
source(here("Scripts", "DataAppendixScripts", "cache_loader.R"))

cache_load_existing()
```

## Main misspecification simulation

Simulate data from the ideal population model and fit the analysis CFA model across the sample-size grid in SEQ. This yields the reference sampling distribution of fit indices under the correctly specified condition.

```{r}
#| label: sim-ideal-popmodel
#| cache: true

popmodel_ideal <- sim(
    NULL,
    n = SEQ,
    analyzemodel,
    generate = popmodel,
    std.lv = TRUE,
    lavaanfun = "cfa",
    seed = SEED,
    multicore = TRUE
)
```

Simulate data from the alternative population model and fit the same analysis CFA model across the same sample-size grid. This creates the sampling distribution of fit indices under misspecification, which is later contrasted with the reference condition to evaluate power.

```{r}
#| label: sim-ideal-h1model
#| cache: true

h1model_ideal <- sim(
    NULL,
    n = SEQ,
    analyzemodel,
    generate = h1model,
    std.lv = TRUE,
    lavaanfun = "cfa",
    seed = SEED,
    multicore = TRUE
)
```

## Main misspecification outputs

### Power curves for fit indices

The following figure summarizes power for rejecting the misspecified condition using the selected fit indices. In `simsem`, `plotPowerFit()` visualizes how this power changes as a function of sample size, using the alternative-model results against the reference-model results and the chosen significance level \(\alpha\) to determine the power at each sample size. The curves show the probability of correctly rejecting the null hypothesis of good fit (i.e., accepting the misspecified model) across different sample sizes, which helps to identify the sample size needed to achieve a desired level of power (e.g., 80%) for detecting misspecification using fit indices.

```{r}
#| label: fig-power-curve-ideal
#| fig-cap: "Power curves for fit indices under ideal conditions."
#| fig-width: 8
#| fig-height: 5

plotPowerFit(
    h1model_ideal,
    popmodel_ideal,
    alpha = ALPHA,
    usedFit = FITS
)
```

The following figure summarizes empirical fit-index distributions for the misspecified condition against the reference condition across the sample-size sequence in N: `usedFit` restricts the display to the selected indices in `FITS` and `logistic = FALSE` requests the raw overlaid scatterplot with cutoff lines instead of logistic-smoothed power curves.

```{r}
#| label: fig-fit-curve-ideal
#| fig-cap: "Power curves for fit indices under ideal conditions without logistic smoothing."
#| fig-width: 8
#| fig-height: 5

plotPowerFit(
    h1model_ideal,
    popmodel_ideal,
    alpha = ALPHA,
    usedFit = FITS,
    logistic = FALSE
)
```

### Power cutoff values for fit indices

This chunk estimates, at \(N = 231\), the empirical power to reject the misspecified model using selected fit indices and, in parallel, derives the corresponding cutoff values from the null model under the same \(\alpha\) level and sample size. In this setup, power is the proportion of replications in which the misspecified model shows worse fit than the cutoff implied by the null-model sampling distribution, which aligns with the general SEM view of power as the probability of rejecting a false null hypothesis [@jobst2023; @moshagen2023].

The logic is straightforward: `getPowerFit()` evaluates how often the alternative model would be rejected, whereas `getCutoff()` obtains index-specific critical values from the null model. Using empirically derived cutoffs is preferable to generic rules of thumb when the goal is to tailor the decision criterion to the specific model, sample size, and analysis conditions [@jobst2023; @moshagen2023].

```{r}
#| label: tbl-power-cutoff-ideal
#| tbl-cap: "Power and fit index cutoff values under ideal conditions at N = 231."

# Estimate power for rejecting the misspecified model (h1model_ideal)
# using fit-index cutoffs implied by the null model at alpha = ALPHA and N = 231.
power_ideal <- getPowerFit(
    h1model_ideal,
    nullObject = popmodel_ideal,
    alpha = ALPHA,
    nVal = 231,
    usedFit = FITS
)

# Derive empirical cutoff values for the selected fit indices
# from the null-model sampling distribution under the same conditions.
cutoff_ideal <- getCutoff(
    popmodel_ideal,
    alpha = ALPHA,
    nVal = 231,
    usedFit = FITS
)

# Combine power and cutoff values into a compact summary table.
power_cutoff_df <- rbind(
    "Power" = as.numeric(power_ideal),
    "Cutoff" = as.numeric(cutoff_ideal[1, ])
)

colnames(power_cutoff_df) <- toupper(names(power_ideal)) # Standardize column names for presentation.


knitr::kable(power_cutoff_df, digits = 3, align = "c") # Print the table.
```

### Tradidional cutoff values for fit indices

```{r}
#| label: fig-fit-curve-traditional
#| fig-cap: "Power curves using traditional fit index cutoffs under ideal conditions."
#| fig-width: 8
#| fig-height: 5

plotPowerFit(
    h1model_ideal,
    popmodel_ideal,
    cutoff = RULE_OF_THUMB,
    alpha = ALPHA,
    usedFit = FITS,
    logistic = FALSE
)
```


```{r}
#| label: tbl-power-cutoff-traditional
#| tbl-cap: "Power for traditional fit index cutoffs under ideal conditions at N = 231."

getPowerFit(
    h1model_ideal,
    RULE_OF_THUMB,
    alpha = ALPHA,
    nVal = 231,
    usedFit = FITS
)
```

```{r}
#| label: tbl-fit-cutoff-traditional
#| tbl-cap: "Traditional fit index cutoff values under ideal conditions."

summaryFit(h1model_ideal, alpha = ALPHA, usedFit = FITS)
```

### Target parameter estimates


```{r}
#| label: tbl-param-estimate-ideal
#| tbl-cap: "Summary of parameter estimates under ideal conditions."

summaryParam(h1model_ideal, alpha = ALPHA, digits = 3, detail = TRUE)
```

```{r}
#| label: tbl-power-n-ideal
#| tbl-cap: "Sample size needed for 80% power under ideal conditions."

pow_h1 <- getPower(h1model_ideal, alpha = ALPHA)
findPower(pow_h1, "N", 0.80)
```

### Recovering the population parameters

Recovery of population parameters is not best strategy in CFA power analyses. The idea is to determine the sample size needed to accurately estimate the population parameters (e.g., factor loadings <> 0) with a certain level of precision. This can be done by simulating data under different models and evaluating how well the analysis model recovers the population parameters. 

But this strategy its not pratical because uselly we don't need an large sample size to recover the loadings > 0, but rather to detect misspecifications (e.g., cross-loadings) that are not present in the population model. Therefore, we can use the power to detect misspecifications as a more relevant criterion for determining the required sample size.

Naive model: tau-equivalent loadings (uniform λ per domain) with same reliability as the population model. This is a common but often unrealistic assumption in CFA power analyses.

```{r}
#| label: sim-ideal-naivemodel
#| cache: true

naivemodel_ideal <- sim(
    NULL,
    n = SEQ,
    analyzemodel,
    generate = naivemodel,
    std.lv = TRUE,
    lavaanfun = "cfa",
    seed = SEED,
    multicore = TRUE
)
```

Maybe the better strategy is recovering N from the lowest population loadings.

```{r}
#| label: tbl-power-n-naive
#| tbl-cap: "Sample size needed for 80% power under naive model."

pow_naive <- getPower(naivemodel_ideal, alpha = ALPHA)
findPower(pow_naive, "N", 0.80)
```

Optimal model: tau-equivalent loadings (uniform λ per domain) with same reliability as the population model, but with higher loadings to match the population reliability. This represents an idealized scenario where the analysis model is correctly specified in terms of reliability, but still assumes tau-equivalence.

```{r}
#| label: sim-ideal-optmodel
#| cache: true

optmodel_ideal <- sim(
    NULL,
    n = SEQ,
    analyzemodel,
    generate = optmodel,
    std.lv = TRUE,
    lavaanfun = "cfa",
    seed = SEED,
    multicore = TRUE
)
```

Maybe the better strategy is recovering N from the lowest population loadings.

```{r}
#| label: tbl-power-n-optimal
#| tbl-cap: "Sample size needed for 80% power under optimal model."

pow_opt <- getPower(optmodel_ideal, alpha = ALPHA)
findPower(pow_opt, "N", 0.80)
```

### Misspecification simulation with different population parameters

#### Naive model

- Power curves for fit indices

```{r}
#| label: fig-power-curve-naive
#| fig-cap: "Power curves for fit indices using the naive population model."
#| fig-width: 8
#| fig-height: 5

plotPowerFit(
    h1model_ideal,
    naivemodel_ideal,
    alpha = ALPHA,
    usedFit = FITS
)
```


```{r}
#| label: fig-fit-curve-naive
#| fig-cap: "Power curves for fit indices using the naive population model without logistic smoothing."
#| fig-width: 8
#| fig-height: 5

plotPowerFit(
    h1model_ideal,
    naivemodel_ideal,
    alpha = ALPHA,
    usedFit = FITS,
    logistic = FALSE
)
```

- Power cutoff values for fit indices

```{r}
#| label: tbl-power-cutoff-naive
#| tbl-cap: "Power and fit index cutoff values using the naive population model at N = 330."

power_naive <- getPowerFit(
    h1model_ideal,
    nullObject = naivemodel_ideal,
    alpha = ALPHA,
    nVal = 330,
    usedFit = FITS
)

cutoff_naive <- getCutoff(
    naivemodel_ideal,
    alpha = ALPHA,
    nVal = 330,
    usedFit = FITS
)

power_cutoff_naive_df <- rbind(
    "Power" = as.numeric(power_naive),
    "Cutoff" = as.numeric(cutoff_naive[1, ])
)
colnames(power_cutoff_naive_df) <- toupper(names(power_naive))

knitr::kable(power_cutoff_naive_df, digits = 3, align = "c")
```

#### Optimal model

- Power curves for fit indices

```{r}
#| label: fig-power-curve-opt
#| fig-cap: "Power curves for fit indices using the optimal population model."
#| fig-width: 8
#| fig-height: 5

plotPowerFit(
    h1model_ideal,
    optmodel_ideal,
    alpha = ALPHA,
    usedFit = FITS
)
```


```{r}
#| label: fig-fit-curve-opt
#| fig-cap: "Power curves for fit indices using the optimal population model without logistic smoothing."
#| fig-width: 8
#| fig-height: 5

plotPowerFit(
    h1model_ideal,
    optmodel_ideal,
    alpha = ALPHA,
    usedFit = FITS,
    logistic = FALSE
)
```

- Power cutoff values for fit indices

```{r}
#| label: tbl-power-cutoff-opt
#| tbl-cap: "Power and fit index cutoff values using the optimal population model at N = 291."

power_opt <- getPowerFit(
    h1model_ideal,
    nullObject = optmodel_ideal,
    alpha = ALPHA,
    nVal = 291,
    usedFit = FITS
)

cutoff_opt <- getCutoff(
    optmodel_ideal,
    alpha = ALPHA,
    nVal = 291,
    usedFit = FITS
)

power_cutoff_opt_df <- rbind(
    "Power" = as.numeric(power_opt),
    "Cutoff" = as.numeric(cutoff_opt[1, ])
)
colnames(power_cutoff_opt_df) <- toupper(names(power_opt))

knitr::kable(power_cutoff_opt_df, digits = 3, align = "c")
```
