---
title: "Simulation — Realistic Scenarios"
execute:
  echo: false
---

## Setup

```{r}
#| label: setup
#| include: false

library(here)
source(here("Scripts", "DataAppendixScripts", "models.R"))
source(here("Scripts", "DataAppendixScripts", "helpers.R"))
source(here("Scripts", "DataAppendixScripts", "cache_utils.R"))
source(here("Scripts", "DataAppendixScripts", "cache_loader.R"))

cache_load_existing()
```

## Scenario A — Missing data (MCAR 10%)

Introduce 10% missing completeness at random (MCAR). This demonstrates the
power cost of attrition / item non-response. For more complex missing data patterns (e.g., MAR, MNAR), you would need to specify a more detailed missing data mechanism (`miss` function in `simsem` package), but this serves as a baseline for the impact of missingness on power. We use FIML estimation to handle the missing data, which is a common approach in SEM.

```{r}
#| label: sim-missing-popmodel
#| cache: true

popmodel_miss <- cache_or_run("popmodel_miss", {
    sim(
        NULL,
        n = SEQ1,              # See the `SEQ1` object defined in helpers.R for the sample size sequence used in this scenario.
        pmMCAR = 0.10,
        analyzemodel,
        generate = popmodel,
        std.lv = TRUE,
        lavaanfun = "cfa",
        estimator = "ml",
        missing = "fiml",
        seed = SEED,
        multicore = TRUE
    )
})
```


```{r}
#| label: sim-missing-h1model
#| cache: true

h1model_miss <- cache_or_run("h1model_miss", {
    sim(
        NULL,
        n = SEQ1,              # See the `SEQ1` object defined in helpers.R for the sample size sequence used in this scenario.
        pmMCAR = 0.10,
        analyzemodel,
        generate = h1model,
        std.lv = TRUE,
        lavaanfun = "cfa",
        estimator = "ml",
        missing = "fiml",
        seed = SEED,
        multicore = TRUE
    )
})
```

### Power curves for fit indices

These chunks summarize a simulation-based, a priori power analysis for global fit under $10\%$ MCAR missingness. The goal is to evaluate how much sample size is needed for the analysis model to reliably distinguish itself from a misspecified competing model when the decision is based on RMSEA, SRMR, CFI, and TLI, following the logic of model-based Monte Carlo power analysis in SEM [@moshagen2024; @feng2023].

```{r}
#| label: fig-power-curve-missing
#| fig-cap: "Power curves for fit indices under missing data conditions."
#| fig-width: 8
#| fig-height: 5

# Plot model-based power as a function of N.
# The null object provides the reference cutoff distribution,
# and the alternative object represents the misspecified model.
# With logistic = TRUE, simsem smooths the rejection probability
# over N using logistic regression, which is useful for identifying
# the sample size at which power approaches the target level.

plotPowerFit(
    h1model_miss,
    popmodel_miss,
    alpha = ALPHA,
    usedFit = FITS
)
```

```{r}
#| label: fig-fit-curve-missing
#| fig-cap: "Power curves for fit indices under missing data conditions without logistic smoothing."
#| fig-width: 8
#| fig-height: 5

# Plot the same power information without logistic smoothing.
# Here the figure shows the simulated fit-index values directly across N,
# which helps visualize how the sampling distributions separate as sample size increases @simsem.

plotPowerFit(
    h1model_miss,
    popmodel_miss,
    alpha = ALPHA,
    usedFit = FITS,
    logistic = FALSE
)
```

### Power cutoff values for fit indices

The next chunk summarizes the model-based decision quantities used in the power analysis. In `simsem`, power is obtained by comparing the sampling distribution under the misspecified model against cutoffs derived from the reference model at the chosen $\alpha$ level, for the same sample size and missing-data condition [@simsem2021; @jobst2023].

```{r}
#| label: tbl-power-cutoff-missing
#| tbl-cap: "Power and fit index cutoff values under missing data conditions at N = 359."

# Estimate power for rejecting the misspecified model
# using fit-index cutoffs implied by the reference model at alpha = .05 and N = 359.
power_miss <- getPowerFit(
    h1model_miss,
    nullObject = popmodel_miss,
    alpha = ALPHA,
    nVal = 359,
    usedFit = FITS
)

# Derive empirical cutoff values for the selected fit indices
# from the reference-model sampling distribution under the same conditions.
cutoff_miss <- getCutoff(
    popmodel_miss,
    alpha = ALPHA,
    nVal = 359,
    usedFit = FITS
)

# Combine power and cutoff values into a compact summary table.
power_cutoff_miss <- rbind(
    "Power" = as.numeric(power_miss),
    "Cutoff" = as.numeric(cutoff_miss[1, ])
)

colnames(power_cutoff_miss) <- toupper(names(power_miss)) # Standardize column names for presentation.

knitr::kable(power_cutoff_miss, digits = 3, align = "c") # Print the table.
```


## Scenario B — Non-normality

This scenario simulates data with varying degrees of non-normality by applying marginal transformations to the observed variables. The `dist` object defined in `helpers.R` specifies a range of skewness and kurtosis values to create different non-normal conditions. This allows us to evaluate how robust fit indices perform when the normality assumption is violated, which is common in real-world data. We use the same estimation method (MLR) that provides robust standard errors and fit indices under non-normal conditions.

```{r}
#| label: sim-nnorm-popmodel
#| cache: true

popmodel_nnorm <- cache_or_run("popmodel_nnorm", {
    sim(
        NULL,
        n = SEQ1,             # See the `SEQ1` object defined in helpers.R for the sample size sequence used in this scenario.
        analyzemodel,
        generate = popmodel,
        indDist = dist,       # See the `dist` object defined in helpers.R for the range of skewness and kurtosis values.
        std.lv = TRUE,
        lavaanfun = "cfa",
        estimator = "mlr",
        seed = SEED,
        multicore = TRUE
    )
})
```


```{r}
#| label: sim-nnorm-h1model
#| cache: true

h1model_nnorm <- cache_or_run("h1model_nnorm", {
    sim(
        NULL,
        n = SEQ1,             # See the `SEQ1` object defined in helpers.R for the sample size sequence used in this scenario.
        analyzemodel,
        generate = h1model,
        indDist = dist,       # See the `dist` object defined in helpers.R for the range of skewness and kurtosis values.
        std.lv = TRUE,
        lavaanfun = "cfa",
        estimator = "mlr",
        seed = SEED,
        multicore = TRUE
    )
})
```

### Power curves for fit indices

This section evaluates power to reject a misspecified SEM across varying $N$ using fit-index cutoffs derived from the null-model sampling distribution, which aligns with the simulation-based logic of absolute-fit power analysis in SEM [@simsem2021; @wang2023].

```{r}
#| label: fig-power-curve-nnorm
#| fig-cap: "Power curves for fit indices under non-normal conditions."
#| fig-width: 8
#| fig-height: 5

# Plot power as a function of N for the selected fit indices.
# In simsem, plotPowerFit visualizes rejection power for a misspecified model
# using either cutoffs supplied directly or cutoffs implied by the null model

# With nullObject = popmodel_nnorm, the cutoff for each index is taken from the
# null-model sampling distribution rather than from fixed rules of thumb.

# When a varying condition such as sample size is present and logistic = TRUE
# (the default), the plotted curve is based on logistic regression smoothing
# of the reject/retain decision over that varying condition.

plotPowerFit(
  h1model_nnorm,
  popmodel_nnorm,
  alpha = ALPHA,
  usedFit = FITS
)
```

The next figure shows the same power analysis without logistic smoothing, so the empirical separation between simulated values and the cutoff line can be inspected more directly [@simsem2021]

```{r}
#| label: fig-fit-curve-nnorm
#| fig-cap: "Power curves for fit indices under non-normal conditions without logistic smoothing."
#| fig-width: 8
#| fig-height: 5

# Reproduce the same plot, but disable logistic smoothing.
# In simsem, logistic = FALSE overlays the raw simulated fit-index values and the
# cutoff line, which is useful for checking dispersion and overlap across N

plotPowerFit(
  h1model_nnorm,
  popmodel_nnorm,
  alpha = ALPHA,
  usedFit = FITS,
  logistic = FALSE
)
```

At $N = 228$, the code below extracts both index-specific power and the corresponding empirical cutoff values under the same $\alpha$ level, so the rejection rule and its operating performance are evaluated on the same sampling basis [@simsem2021; @feng2023].

```{r}
#| label: tbl-power-cutoff-nnorm
#| tbl-cap: "Power and fit index cutoff values under non-normal conditions at N = 228."

# Estimate power to reject the misspecified model at N = 228.
# getPowerFit returns the probability that each selected fit index classifies
# the misspecified model as worse than the null-based cutoff
power_nnorm <- getPowerFit(
  h1model_nnorm,
  nullObject = popmodel_nnorm,
  alpha = ALPHA,
  nVal = 228,
  usedFit = FITS
)

# Derive empirical cutoff values from the null-model sampling distribution
# under the same alpha level and sample size
cutoff_nnorm <- getCutoff(
  popmodel_nnorm,
  alpha = ALPHA,
  nVal = 228,
  usedFit = FITS
)

# Combine operating characteristics (power) and decision thresholds (cutoffs)
# into a compact table for direct reporting
power_cutoff_nnorm <- rbind(
  "Power" = as.numeric(power_nnorm),
  "Cutoff" = as.numeric(cutoff_nnorm[1, ])
)

colnames(power_cutoff_nnorm) <- toupper(names(power_nnorm))

knitr::kable(power_cutoff_nnorm, digits = 3, align = "c")
```
