---
title: "Simulation — Realistic Scenarios"
execute:
  echo: false
---

## Setup

```{r}
#| label: setup
#| include: false

library(here)
source(here("Scripts", "DataAppendixScripts", "helpers.R"))
source(here("Scripts", "DataAppendixScripts", "models.R"))
```

## Scenario A — Missing data (MAR 10%)

Introduce 10% missing completeness at random (MCAR). This demonstrates the
power cost of attrition / item non-response. For more complex missing data patterns (e.g., MAR, MNAR), you would need to specify a more detailed missing data mechanism (`miss` function in `simsem` package), but this serves as a baseline for the impact of missingness on power. We use FIML estimation to handle the missing data, which is a common approach in SEM.

```{r}
#| label: sim-missing-popmodel
#| cache: true

popmodel_miss <- sim(
    NULL,
    n = 51:350,
    pmMCAR = 0.10,
    analyzemodel,
    generate = popmodel,
    std.lv = TRUE,
    lavaanfun = "cfa",
    estimator = "ml",
    missing = "fiml",
    seed = SEED,
    multicore = TRUE
)
```


```{r}
#| label: sim-missing-h1model
#| cache: true

h1model_miss <- sim(
    NULL,
    n = 51:350,
    pmMCAR = 0.10,
    analyzemodel,
    generate = h1model,
    std.lv = TRUE,
    lavaanfun = "cfa",
    estimator = "ml",
    missing = "fiml",
    seed = SEED,
    multicore = TRUE
)
```

### Power curves for fit indices

```{r}
#| label: power-curve-missing

plotPowerFit(
    h1model_miss,
    popmodel_miss,
    alpha = ALPHA,
    usedFit = FITS
)
```


```{r}
#| label: fit-curve-missing

plotPowerFit(
    h1model_miss,
    popmodel_miss,
    alpha = ALPHA,
    usedFit = FITS,
    logistic = FALSE
)
```

### Power cutoff values for fit indices

You extract the cutoff values for fit indices at a given alpha level and sample size to determine the threshold for acceptable fit under the null model. This allows you to see how stringent the fit criteria are at different sample sizes.

```{r}
#| label: power-cutoff-missing

getPowerFit(
    h1model_miss,
    nullObject = popmodel_miss,
    alpha = ALPHA,
    nVal = 100,
    usedFit = FITS
)
```

```{r}
#| label: fit-cutoff-missing

getCutoff(
    popmodel_miss,
    alpha = ALPHA,
    nVal = 100,
    usedFit = FITS
)
```


## Scenario B — Non-normality

Generate data from the population model with normal distributions, then
apply marginal transformations to introduce moderate non-normality
(skewness ≈ 1.5, kurtosis ≈ 3). This simulates realistic conditions where
psychological data often deviate from multivariate normality.

We use a custom data generation function that applies a power transformation
to each variable after generation, combined with ML and MLR estimators to
show the impact.


```{r}
#| label: sim-nnorm-popmodel
#| cache: true

dist <- bindDist(
    skewness = seq(-1, 1, length.out = 24), 
    kurtosis = seq(1, 2, length.out = 24)
)

popmodel_nnorm <- sim(
    NULL,
    n = 51:350,
    analyzemodel,
    generate = popmodel,
    indDist = dist,
    std.lv = TRUE,
    lavaanfun = "cfa",
    estimator = "mlr",
    seed = SEED,
    multicore = TRUE
)
```


```{r}
#| label: sim-nnorm-h1model
#| cache: true

h1model_nnorm <- sim(
    NULL,
    n = 51:350,
    analyzemodel,
    generate = h1model,
    indDist = dist,
    std.lv = TRUE,
    lavaanfun = "cfa",
    estimator = "mlr",
    seed = SEED,
    multicore = TRUE
)
```

### Power curves for fit indices

```{r}
#| label: power-curve-nnorm

plotPowerFit(
    h1model_nnorm,
    popmodel_nnorm,
    alpha = ALPHA,
    usedFit = FITS_ROB
)
```


```{r}
#| label: fit-curve-nnorm

plotPowerFit(
    h1model_nnorm,
    popmodel_nnorm,
    alpha = ALPHA,
    usedFit = FITS_ROB,
    logistic = FALSE
)
```

### Power cutoff values for fit indices

You extract the cutoff values for fit indices at a given alpha level and sample size to determine the threshold for acceptable fit under the null model. This allows you to see how stringent the fit criteria are at different sample sizes.

```{r}
#| label: power-cutoff-nnorm

getPowerFit(
    h1model_nnorm,
    nullObject = popmodel_nnorm,
    alpha = ALPHA,
    nVal = 100,
    usedFit = FITS_ROB
)
```

```{r}
#| label: fit-cutoff-nnorm

getCutoff(
    popmodel_nnorm,
    alpha = ALPHA,
    nVal = 100,
    usedFit = FITS_ROB
)
```


## Scenario C — Missing data + Non-normality

This scenario combines the challenges of missing data and non-normality, reflecting common issues in real-world datasets. We generate data with 10% MAR missingness and apply the same marginal transformations to introduce non-normality.

```{r}
#| label: sim-normiss-popmodel
#| cache: true

popmodel_normiss <- sim(
    NULL,
    n = 51:350,
    pmMCAR = 0.10,
    analyzemodel,
    generate = popmodel,
    indDist = dist,
    std.lv = TRUE,
    lavaanfun = "cfa",
    estimator = "mlr",
    missing = "fiml",
    seed = SEED,
    multicore = TRUE
)
```


```{r}
#| label: sim-normiss-h1model
#| cache: true

h1model_normiss <- sim(
    NULL,
    n = 51:350,
    pmMCAR = 0.10,
    analyzemodel,
    generate = h1model,
    indDist = dist,
    std.lv = TRUE,
    lavaanfun = "cfa",
    estimator = "mlr",
    missing = "fiml",
    seed = SEED,
    multicore = TRUE
)
```

### Power curves for fit indices

```{r}
#| label: power-curve-normiss

plotPowerFit(
    h1model_normiss,
    popmodel_normiss,
    alpha = ALPHA,
    usedFit = FITS_ROB
)
```


```{r}
#| label: fit-curve-normiss

plotPowerFit(
    h1model_normiss,
    popmodel_normiss,
    alpha = ALPHA,
    usedFit = FITS_ROB,
    logistic = FALSE
)
```


### Power cutoff values for fit indices

You extract the cutoff values for fit indices at a given alpha level and sample size to determine the threshold for acceptable fit under the null model. This allows you to see how stringent the fit criteria are at different sample sizes.

```{r}
#| label: power-cutoff-normiss

getPowerFit(
    h1model_normiss,
    nullObject = popmodel_normiss,
    alpha = ALPHA,
    nVal = 135,
    usedFit = FITS_ROB
)
```

```{r}
#| label: fit-cutoff-normiss

getCutoff(
    popmodel_normiss,
    alpha = ALPHA,
    nVal = 135,
    usedFit = FITS_ROB
)
```
