---
title: "Simulation — Realistic Scenarios"
execute:
  echo: false
  warning: false
  message: false
---

## Setup

```{r}
#| label: setup
#| include: false

library(here)
source(here("Scripts", "DataAppendixScripts", "models.R"))
source(here("Scripts", "DataAppendixScripts", "helpers.R"))
source(here("Scripts", "DataAppendixScripts", "cache_utils.R"))
source(here("Scripts", "DataAppendixScripts", "cache_loader.R"))

cache_load_existing()
```

## Scenario A — Missing data (MCAR 10%)

Missingness reduces effective sample size and inflates sampling variance,
both of which reduce power relative to the ideal scenario. When data are
missing completely at random (MCAR), this reduction is a function of the
missing proportion alone, and Full Information Maximum Likelihood (FIML)
provides consistent estimates without imputation [@enders2010; @simsem2021].

The two chunks below replicate the main misspecification simulation with
10% MCAR applied to both the reference and the alternative conditions.
Using the same population and analysis models as before isolates the power
cost of missingness, since all other design choices remain constant.

```{r}
#| label: sim-missing-popmodel
#| cache: true

popmodel_miss <- cache_or_run("popmodel_miss", {
    sim(
        NULL,
        n = SEQ1,              # See the `SEQ1` object defined in helpers.R for the sample size sequence used in this scenario.
        pmMCAR = 0.10,
        analyzemodel,
        generate = popmodel,
        std.lv = TRUE,
        lavaanfun = "cfa",
        estimator = "ml",
        missing = "fiml",
        seed = SEED,
        multicore = TRUE
    )
})
```

```{r}
#| label: sim-missing-h1model
#| cache: true

h1model_miss <- cache_or_run("h1model_miss", {
    sim(
        NULL,
        n = SEQ1,              # See the `SEQ1` object defined in helpers.R for the sample size sequence used in this scenario.
        pmMCAR = 0.10,
        analyzemodel,
        generate = h1model,
        std.lv = TRUE,
        lavaanfun = "cfa",
        estimator = "ml",
        missing = "fiml",
        seed = SEED,
        multicore = TRUE
    )
})
```

### Power curves for fit indices

The power curves below follow the same logic as in the ideal scenario,
but with cutoff distributions now derived from samples subject to 10%
MCAR. The expected outcome is a rightward shift in the curves: the same
target power level requires a larger $N$ than under complete data
[@enders2010; @feng2023].

```{r}
#| label: fig-power-curve-missing
#| fig-cap: "Power curves for fit indices under missing data conditions."
#| fig-width: 8
#| fig-height: 5

# Plot model-based power as a function of N.
# The null object provides the reference cutoff distribution,
# and the alternative object represents the misspecified model.
# With logistic = TRUE, simsem smooths the rejection probability
# over N using logistic regression, which is useful for identifying
# the sample size at which power approaches the target level.

plotPowerFit(
    h1model_miss,
    popmodel_miss,
    alpha = ALPHA,
    usedFit = FITS
)
```

```{r}
#| label: fig-fit-curve-missing
#| fig-cap: "Power curves for fit indices under missing data conditions without logistic smoothing."
#| fig-width: 8
#| fig-height: 5

# Plot the same power information without logistic smoothing.
# Here the figure shows the simulated fit-index values directly across N,
# which helps visualize how the sampling distributions separate as sample size increases @simsem.

plotPowerFit(
    h1model_miss,
    popmodel_miss,
    alpha = ALPHA,
    usedFit = FITS,
    logistic = FALSE
)
```

### Power and cutoff values for fit indices

`getPowerFit()` and `getCutoff()` are evaluated here at $N = 359$,
the minimum sample size at which all four indices reached the target
power under this missing-data condition. Reading power alongside the
empirical cutoffs makes it possible to verify that the decision rule
itself is well-calibrated — not merely that the rejection rate is
high at the chosen $N$ [@jobst2023; @wang2023].

```{r}
#| label: tbl-power-cutoff-missing
#| tbl-cap: "Power and fit index cutoff values under missing data conditions at N = 359."

# Estimate power for rejecting the misspecified model
# using fit-index cutoffs implied by the reference model at alpha = .05 and N = 359.
power_miss <- getPowerFit(
    h1model_miss,
    nullObject = popmodel_miss,
    alpha = ALPHA,
    nVal = 359,
    usedFit = FITS
)

# Derive empirical cutoff values for the selected fit indices
# from the reference-model sampling distribution under the same conditions.
cutoff_miss <- getCutoff(
    popmodel_miss,
    alpha = ALPHA,
    nVal = 359,
    usedFit = FITS
)

# Combine power and cutoff values into a compact summary table.
power_cutoff_miss <- rbind(
    "Power" = as.numeric(power_miss),
    "Cutoff" = as.numeric(cutoff_miss[1, ])
)

colnames(power_cutoff_miss) <- toupper(names(power_miss)) # Standardize column names for presentation.

knitr::kable(power_cutoff_miss, digits = 3, align = "c") # Print the table.
```

## Scenario B — Non-normality

Non-normal continuous indicators are common in psychometric research,
particularly when items use ordered response scales or when the
construct exhibits floor or ceiling effects.
Under non-normality, standard ML standard errors and test statistics
are generally biased, so Satorra-Bentler scaling (MLR) is applied here
to maintain appropriate Type I error control.

The `dist` object in `helpers.R` generates data with mild skewness
(ranging from $-1$ to $+1$) and slight excess kurtosis (from $1$ to
$2$), a range that is realistic for quality-of-life scales such as
the WHOQOL-BREF. The goal is not to impose severe
distributional violations, but to verify that the power conclusions
from the ideal scenario hold under plausible real-world deviations
from normality.

```{r}
#| label: sim-nnorm-popmodel
#| cache: true

popmodel_nnorm <- cache_or_run("popmodel_nnorm", {
    sim(
        NULL,
        n = SEQ1,             # See the `SEQ1` object defined in helpers.R for the sample size sequence used in this scenario.
        analyzemodel,
        generate = popmodel,
        indDist = dist,       # See the `dist` object defined in helpers.R for the range of skewness and kurtosis values.
        std.lv = TRUE,
        lavaanfun = "cfa",
        estimator = "mlr",
        seed = SEED,
        multicore = TRUE
    )
})
```

```{r}
#| label: sim-nnorm-h1model
#| cache: true

h1model_nnorm <- cache_or_run("h1model_nnorm", {
    sim(
        NULL,
        n = SEQ1,             # See the `SEQ1` object defined in helpers.R for the sample size sequence used in this scenario.
        analyzemodel,
        generate = h1model,
        indDist = dist,       # See the `dist` object defined in helpers.R for the range of skewness and kurtosis values.
        std.lv = TRUE,
        lavaanfun = "cfa",
        estimator = "mlr",
        seed = SEED,
        multicore = TRUE
    )
})
```

### Power curves for fit indices

As in the ideal scenario, power is evaluated against empirical cutoffs
derived from the null-model sampling distribution, now obtained under
MLR estimation. This keeps the decision criterion consistent with the
data conditions and avoids applying cutoffs calibrated under normality
to non-normal simulated data [@mcneish2021; @simsem2021].

```{r}
#| label: fig-power-curve-nnorm
#| fig-cap: "Power curves for fit indices under non-normal conditions."
#| fig-width: 8
#| fig-height: 5

# Plot power as a function of N for the selected fit indices.
# In simsem, plotPowerFit visualizes rejection power for a misspecified model
# using either cutoffs supplied directly or cutoffs implied by the null model

# With nullObject = popmodel_nnorm, the cutoff for each index is taken from the
# null-model sampling distribution rather than from fixed rules of thumb.

# When a varying condition such as sample size is present and logistic = TRUE
# (the default), the plotted curve is based on logistic regression smoothing
# of the reject/retain decision over that varying condition.

plotPowerFit(
  h1model_nnorm,
  popmodel_nnorm,
  alpha = ALPHA,
  usedFit = FITS
)
```

The unsmoothed version below displays the raw separation between
reference and alternative fit-index distributions across sample sizes,
which helps assess at what point the two conditions become
empirically distinguishable [@simsem2021; @mcneish2021].

```{r}
#| label: fig-fit-curve-nnorm
#| fig-cap: "Power curves for fit indices under non-normal conditions without logistic smoothing."
#| fig-width: 8
#| fig-height: 5

# Reproduce the same plot, but disable logistic smoothing.
# In simsem, logistic = FALSE overlays the raw simulated fit-index values and the
# cutoff line, which is useful for checking dispersion and overlap across N

plotPowerFit(
  h1model_nnorm,
  popmodel_nnorm,
  alpha = ALPHA,
  usedFit = FITS,
  logistic = FALSE
)
```

### Power and cutoff values for fit indices

At $N = 228$, the table below summarizes power and empirical cutoffs
under mild non-normality. Comparing these values with the ideal-scenario
results at $N = 231$ shows the net effect of the distributional
deviation: if power and cutoff values remain close across the two
conditions, the conclusions from the ideal scenario are robust to
this degree of non-normality [@feng2023].

```{r}
#| label: tbl-power-cutoff-nnorm
#| tbl-cap: "Power and fit index cutoff values under non-normal conditions at N = 228."

# Estimate power to reject the misspecified model at N = 228.
# getPowerFit returns the probability that each selected fit index classifies
# the misspecified model as worse than the null-based cutoff
power_nnorm <- getPowerFit(
  h1model_nnorm,
  nullObject = popmodel_nnorm,
  alpha = ALPHA,
  nVal = 228,
  usedFit = FITS
)

# Derive empirical cutoff values from the null-model sampling distribution
# under the same alpha level and sample size
cutoff_nnorm <- getCutoff(
  popmodel_nnorm,
  alpha = ALPHA,
  nVal = 228,
  usedFit = FITS
)

# Combine operating characteristics (power) and decision thresholds (cutoffs)
# into a compact table for direct reporting
power_cutoff_nnorm <- rbind(
  "Power" = as.numeric(power_nnorm),
  "Cutoff" = as.numeric(cutoff_nnorm[1, ])
)

colnames(power_cutoff_nnorm) <- toupper(names(power_nnorm))

knitr::kable(power_cutoff_nnorm, digits = 3, align = "c")
```
