---
title: "Simulation — Realistic Scenarios"
execute:
  echo: false
---

## Setup

```{r}
#| label: setup
#| include: false

library(here)
source(here("Scripts", "DataAppendixScripts", "helpers.R"))
source(here("Scripts", "DataAppendixScripts", "models.R"))
```

## Scenario A — Missing data (MCAR 10%)

Introduce 10% missing completeness at random (MCAR). This demonstrates the
power cost of attrition / item non-response. For more complex missing data patterns (e.g., MAR, MNAR), you would need to specify a more detailed missing data mechanism (`miss` function in `simsem` package), but this serves as a baseline for the impact of missingness on power. We use FIML estimation to handle the missing data, which is a common approach in SEM.

```{r}
#| label: sim-missing-popmodel
#| cache: true

popmodel_miss <- sim(
    NULL,
    n = SEQ,
    pmMCAR = 0.10,
    analyzemodel,
    generate = popmodel,
    std.lv = TRUE,
    lavaanfun = "cfa",
    estimator = "ml",
    missing = "fiml",
    seed = SEED,
    multicore = TRUE
)
```


```{r}
#| label: sim-missing-h1model
#| cache: true

h1model_miss <- sim(
    NULL,
    n = SEQ,
    pmMCAR = 0.10,
    analyzemodel,
    generate = h1model,
    std.lv = TRUE,
    lavaanfun = "cfa",
    estimator = "ml",
    missing = "fiml",
    seed = SEED,
    multicore = TRUE
)
```

### Power curves for fit indices

```{r}
#| label: power-curve-missing
#| fig-cap: "Power curves for fit indices under missing data conditions."
#| fig-width: 8
#| fig-height: 5

plotPowerFit(
    h1model_miss,
    popmodel_miss,
    alpha = ALPHA,
    usedFit = FITS
)
```


```{r}
#| label: fit-curve-missing
#| fig-cap: "Power curves for fit indices under missing data conditions without logistic smoothing."
#| fig-width: 8
#| fig-height: 5

plotPowerFit(
    h1model_miss,
    popmodel_miss,
    alpha = ALPHA,
    usedFit = FITS,
    logistic = FALSE
)
```

### Power cutoff values for fit indices

You extract the cutoff values for fit indices at a given alpha level and sample size to determine the threshold for acceptable fit under the null model. This allows you to see how stringent the fit criteria are at different sample sizes.

```{r}
#| label: power-cutoff-missing
#| tbl-cap: "Power for fit index cutoffs under missing data conditions at N = 297."

getPowerFit(
    h1model_miss,
    nullObject = popmodel_miss,
    alpha = ALPHA,
    nVal = 297,
    usedFit = FITS
)
```

```{r}
#| label: fit-cutoff-missing
#| tbl-cap: "Fit index cutoff values under missing data conditions at N = 100."

getCutoff(
    popmodel_miss,
    alpha = ALPHA,
    nVal = 297,
    usedFit = FITS
)
```


## Scenario B — Non-normality

Generate data from the population model with normal distributions, then
apply marginal transformations to introduce moderate non-normality
(skewness ≈ 1.5, kurtosis ≈ 3). This simulates realistic conditions where
psychological data often deviate from multivariate normality.

We use a custom data generation function that applies a power transformation
to each variable after generation, combined with ML and MLR estimators to
show the impact.


```{r}
#| label: sim-nnorm-popmodel
#| cache: true

dist <- bindDist(
    skewness = seq(-1, 1, length.out = 24), 
    kurtosis = seq(1, 2, length.out = 24)
)

popmodel_nnorm <- sim(
    NULL,
    n = SEQ,
    analyzemodel,
    generate = popmodel,
    indDist = dist,
    std.lv = TRUE,
    lavaanfun = "cfa",
    estimator = "mlr",
    seed = SEED,
    multicore = TRUE
)
```


```{r}
#| label: sim-nnorm-h1model
#| cache: true

h1model_nnorm <- sim(
    NULL,
    n = SEQ,
    analyzemodel,
    generate = h1model,
    indDist = dist,
    std.lv = TRUE,
    lavaanfun = "cfa",
    estimator = "mlr",
    seed = SEED,
    multicore = TRUE
)
```

### Power curves for fit indices

```{r}
#| label: power-curve-nnorm
#| fig-cap: "Power curves for robust fit indices under non-normal conditions."
#| fig-width: 8
#| fig-height: 5

plotPowerFit(
    h1model_nnorm,
    popmodel_nnorm,
    alpha = ALPHA,
    usedFit = FITS_ROB
)
```


```{r}
#| label: fit-curve-nnorm
#| fig-cap: "Power curves for robust fit indices under non-normal conditions without logistic smoothing."
#| fig-width: 8
#| fig-height: 5

plotPowerFit(
    h1model_nnorm,
    popmodel_nnorm,
    alpha = ALPHA,
    usedFit = FITS_ROB,
    logistic = FALSE
)
```

### Power cutoff values for fit indices

You extract the cutoff values for fit indices at a given alpha level and sample size to determine the threshold for acceptable fit under the null model. This allows you to see how stringent the fit criteria are at different sample sizes.

```{r}
#| label: power-cutoff-nnorm
#| tbl-cap: "Power for robust fit index cutoffs under non-normal conditions at N = 201."

getPowerFit(
    h1model_nnorm,
    nullObject = popmodel_nnorm,
    alpha = ALPHA,
    nVal = 201,
    usedFit = FITS_ROB
)
```

```{r}
#| label: fit-cutoff-nnorm
#| tbl-cap: "Robust fit index cutoff values under non-normal conditions at N = 201."

getCutoff(
    popmodel_nnorm,
    alpha = ALPHA,
    nVal = 201,
    usedFit = FITS_ROB
)
```


## Scenario C — Missing data + Non-normality

This scenario combines the challenges of missing data and non-normality, reflecting common issues in real-world datasets. We generate data with 10% MAR missingness and apply the same marginal transformations to introduce non-normality.

```{r}
#| label: sim-normiss-popmodel
#| cache: true

popmodel_normiss <- sim(
    NULL,
    n = SEQ,
    pmMCAR = 0.10,
    analyzemodel,
    generate = popmodel,
    indDist = dist,
    std.lv = TRUE,
    lavaanfun = "cfa",
    estimator = "mlr",
    missing = "fiml",
    seed = SEED,
    multicore = TRUE
)
```


```{r}
#| label: sim-normiss-h1model
#| cache: true

h1model_normiss <- sim(
    NULL,
    n = SEQ,
    pmMCAR = 0.10,
    analyzemodel,
    generate = h1model,
    indDist = dist,
    std.lv = TRUE,
    lavaanfun = "cfa",
    estimator = "mlr",
    missing = "fiml",
    seed = SEED,
    multicore = TRUE
)
```

### Power curves for fit indices

```{r}
#| label: power-curve-normiss
#| fig-cap: "Power curves for robust fit indices under missing data and non-normal conditions."
#| fig-width: 8
#| fig-height: 5

plotPowerFit(
    h1model_normiss,
    popmodel_normiss,
    alpha = ALPHA,
    usedFit = FITS_ROB
)
```


```{r}
#| label: fit-curve-normiss
#| fig-cap: "Power curves for robust fit indices under missing data and non-normal conditions without logistic smoothing."
#| fig-width: 8
#| fig-height: 5

plotPowerFit(
    h1model_normiss,
    popmodel_normiss,
    alpha = ALPHA,
    usedFit = FITS_ROB,
    logistic = FALSE
)
```


### Power cutoff values for fit indices

You extract the cutoff values for fit indices at a given alpha level and sample size to determine the threshold for acceptable fit under the null model. This allows you to see how stringent the fit criteria are at different sample sizes.

```{r}
#| label: power-cutoff-normiss
#| tbl-cap: "Power for robust fit index cutoffs under missing data and non-normal conditions at N = 185."

getPowerFit(
    h1model_normiss,
    nullObject = popmodel_normiss,
    alpha = ALPHA,
    nVal = 350,
    usedFit = FITS_ROB
)
```

```{r}
#| label: fit-cutoff-normiss
#| tbl-cap: "Robust fit index cutoff values under missing data and non-normal conditions at N = 185."

getCutoff(
    popmodel_normiss,
    alpha = ALPHA,
    nVal = 350,
    usedFit = FITS_ROB
)
```
