---
title: "Simulation — Fixed N"
execute:
  echo: false
  warning: false
  message: false
---

## Setup

```{r}
#| label: setup
#| include: false

library(here)
source(here("Scripts", "DataAppendixScripts", "models.R"))
source(here("Scripts", "DataAppendixScripts", "helpers.R"))
source(here("Scripts", "DataAppendixScripts", "cache_utils.R"))
source(here("Scripts", "DataAppendixScripts", "cache_loader.R"))

cache_load_existing()
```

## Power analysis with fixed N

The preceding sections determined the minimum $N$ required to achieve
adequate power across a range of conditions. This section takes a
different perspective: rather than searching for the smallest sufficient
sample size, it fixes $N = 350$ and asks how reliably that value
sustains the target power under the most demanding combined scenario —
10% MCAR missingness and mild non-normality simultaneously [@feng2023;
@simsem2021].

This design choice reflects a common applied situation in which the
sample size is constrained by practical or budgetary considerations.
Running a large number of replications at a fixed $N$ provides a direct
stability check: if power remains consistently above the target across
many random samples, the chosen sample size can be considered a robust
planning value for the full research design [@wang2021; @jobst2023].

```{r}
#| label: sim-fixed-popmodel
#| cache: true

popmodel_fixed <- cache_or_run("popmodel_fixed", {
    sim(
        REP,
        n = 350,
        pmMCAR = 0.10,
        analyzemodel,
        generate = popmodel,
        indDist = dist,       # See the `dist` object defined in helpers.R for the range of skewness and kurtosis values.
        std.lv = TRUE,
        lavaanfun = "cfa",
        estimator = "mlr",
        missing = "fiml",
        seed = SEED,
        multicore = TRUE
    )
})
```

```{r}
#| label: sim-fixed-h1model
#| cache: true

h1model_fixed <- cache_or_run("h1model_fixed", {
    sim(
        REP,
        n = 350,
        pmMCAR = 0.10,
        analyzemodel,
        generate = h1model,
        indDist = dist,      # See the `dist` object defined in helpers.R for the range of skewness and kurtosis values.
        std.lv = TRUE,
        lavaanfun = "cfa",
        estimator = "mlr",
        missing = "fiml",
        seed = SEED,
        multicore = TRUE
    )
})
```

### Power curves for fit indices

With $N$ fixed, `plotPowerFit()` no longer traces power as a function
of sample size. Instead, the figure displays the empirical distributions
of each fit index under the reference and misspecified conditions across
replications, alongside the cutoff derived from the reference model at
the chosen $\alpha$ level. The visual separation between the two
distributions is the graphical counterpart of the power values reported
in the table below [@simsem2021; @mcneish2021].

```{r}
#| label: fig-power-curve-fixed
#| fig-cap: "Power curves for fit indices with fixed sample size."
#| fig-width: 8
#| fig-height: 5

plotPowerFit(
    h1model_fixed,
    popmodel_fixed,
    alpha = ALPHA,
    usedFit = FITS
)
```

### Power and cutoff values for fit indices

```{r}
#| label: tbl-power-cutoff-fixed
#| tbl-cap: "Power and fit index cutoff values with fixed sample size."

# Estimate power for rejecting the misspecified model (h1model_fixed)
# using fit-index cutoffs implied by the null model.
power_fixed <- getPowerFit(
  h1model_fixed,
  nullObject = popmodel_fixed,
  alpha = ALPHA,
  usedFit = FITS
)

# Derive empirical cutoff values for the selected fit indices
# from the null-model sampling distribution under the same conditions.
cutoff_fixed <- getCutoff(
  popmodel_fixed,
  alpha = ALPHA,
  usedFit = FITS
)

# Subset power to the four requested fit indices (FITS: rmsea, srmr, cfi, tli)
# ensuring 1-to-1 alignment with cutoff_fixed and excluding TraditionalChi.
power_fixed_clean <- power_fixed[FITS]
cutoff_fixed_clean <- cutoff_fixed[1, FITS]

# Combine power and cutoff values into a compact summary table.
power_cutoff_fixed <- rbind(
  "Power"  = as.numeric(power_fixed_clean),
  "Cutoff" = as.numeric(cutoff_fixed_clean)
)

colnames(power_cutoff_fixed) <- toupper(FITS)

knitr::kable(power_cutoff_fixed, digits = 3, align = "c")
```

## Power analysis with traditional cutoffs

The same fixed-$N$ simulation is now re-evaluated using conventional
rule-of-thumb thresholds in place of the empirically derived cutoffs
used above. This is a deliberate contrast: by holding $N$, the
data-generating model, and the analysis model constant while changing
only the decision criterion, the comparison isolates the contribution
of the cutoff choice to the power conclusion [@mcneish2021;
@kyriazos2018].

```{r}
#| label: fig-curve-fixed-traditional
#| fig-cap: "Power curves using traditional fit index cutoffs with fixed sample size."
#| fig-width: 8
#| fig-height: 5

plotPowerFit(
    h1model_fixed,
    popmodel_fixed,
    cutoff = RULE_OF_THUMB,
    alpha = ALPHA,
    usedFit = FITS
)
```

```{r}
#| label: tbl-power-cutoff-traditional-fixed
#| tbl-cap: "Power for traditional fit index cutoffs with fixed sample size."

# Estimate power using traditional fit-index cutoffs.
power_traditional <- getPowerFit(
  h1model_fixed,
  RULE_OF_THUMB,
  alpha = ALPHA,
  usedFit = FITS
)

# Convert to a one-row table with column headers.
power_traditional_tbl <- t(as.data.frame(power_traditional))

colnames(power_traditional_tbl) <- toupper(names(power_traditional))
rownames(power_traditional_tbl) <- "Power"

knitr::kable(power_traditional_tbl, digits = 3, align = "c")
```


