---
title: "Model Comparison (Null vs Alternative)"
execute:
  echo: true
---

## Setup

```{r}
#| label: setup
#| include: false

library(here)
source(here("Scripts", "DataAppendixScripts", "helpers.R"))
source(here("Scripts", "DataAppendixScripts", "models.R"))
```


## Load empirical data and fit the "null" model

The "null" model here is the empirically estimated CFA model from the
WHOQOL-BREF dataset (Rogers, 2022). The "alternative" is the meta-analytic
population model. The question: can we distinguish between data generated
from the estimated model vs. data generated from the population model?

```{r}
#| label: fit-empirical-model
#| cache: true

# Load WHOQOL-BREF data
url_data <- "https://data.mendeley.com/datasets/rdky78bk8r/2/files/b58a7054-4978-4239-81db-00f9efd86e44/WHOQOL_Data.dat"
url_labels <- "https://data.mendeley.com/datasets/rdky78bk8r/2/files/ae2a5c01-2fc2-43e5-b31c-8db8cc5bff8a/WHOQOL_Labels.txt"

df0 <- read.table(url_data)
column_labels <- read.table(url_labels)$V1
names(df0) <- column_labels

# Strip suffixes from item names (e.g., Q3_F -> Q3)
names(df0) <- gsub("_(P|F|S|A)$", "", names(df0))

# Fit the CFA model to empirical data (continuous, ML for comparability)
myModel <- cfa(
  analyzeModel,
  data   = df0,
  std.lv = TRUE
)

cat("Empirical model fit:\n")
lavaan::fitMeasures(myModel, c("chisq", "df", "pvalue", "cfi", "tli", "rmsea", "srmr"))
```

## Null distribution: Data generated from the fitted model

Simulate data from the estimated (empirical) model. This represents "what
the world looks like if the sample-estimated parameters are the truth."

```{r}
#| label: sim-null-model
#| cache: true

sim_null <- run_sim(
  nRep     = REP,
  n        = N_EMPIRICAL,
  model    = analyzeModel,
  generate = myModel       # generate from the fitted lavaan object
)

cat("Null model simulation summary:\n")
summary(sim_null)
```

## Alternative distribution: Data from the population model

Simulate data from the meta-analytic population model. This represents
"what the world looks like if the meta-analytic population is the truth."

```{r}
#| label: sim-alt-model
#| cache: true

sim_alt <- run_sim(
  nRep     = REP,
  n        = N_EMPIRICAL,
  model    = analyzeModel,
  generate = popModel      # generate from the population model
)

cat("Alternative model simulation summary:\n")
summary(sim_alt)
```

## Power to distinguish null from alternative

Use the cutoffs derived from the null distribution to compute power in the
alternative distribution. This tells us: if the true population follows
the meta-analytic model, can we detect the difference from the empirical
model with N = 1047?

```{r}
#| label: posthoc-comparison

# Get cutoffs from the null (empirical) distribution at alpha = 0.05
cutoff_null <- getCutoff(sim_null, alpha = ALPHA)
cat("Cutoffs from null distribution:\n")
print(cutoff_null)

# Power to detect the alternative using null-derived cutoffs
cat("\nPower (alternative vs. null-derived cutoffs):\n")
getPowerFit(sim_alt, cutoff = cutoff_null)
```

## Visualization: Null vs. Alternative fit distributions

```{r}
#| label: fig-null-vs-alt
#| fig-cap: "Fit index distributions — Null (empirical) vs. Alternative (population)"
#| fig-width: 10
#| fig-height: 8

plotPowerFit(sim_alt, sim_null, alpha = ALPHA)
```

```{r}
#| label: fig-null-vs-alt-selected
#| fig-cap: "Selected fit indices — RMSEA, CFI, SRMR"
#| fig-width: 10
#| fig-height: 6

plotPowerFit(
  sim_alt,
  sim_null,
  alpha   = ALPHA,
  usedFit = c("rmsea", "srmr", "cfi")
)
```

## Summary

```{r}
#| label: tbl-posthoc-fit
#| tbl-cap: "Post hoc power to distinguish empirical from population model (N = 1047)"

posthoc_summary <- data.frame(
  `Fit Index` = c("RMSEA", "CFI", "TLI", "SRMR"),
  `Null Cutoff (α = .05)` = c(
    "From cutoff_null above",
    "From cutoff_null above",
    "From cutoff_null above",
    "From cutoff_null above"
  ),
  `Power` = c(
    "From getPowerFit above",
    "From getPowerFit above",
    "From getPowerFit above",
    "From getPowerFit above"
  ),
  check.names = FALSE
)

kable(posthoc_summary, align = "lcc")
```
