---
title: "Simulation — Fixed N"
execute:
  echo: false
  warning: false
  message: false
---

## Setup

```{r}
#| label: setup
#| include: false

# For defining file paths
library(here)

# For CFA population models
source(here("Scripts", "DataAppendixScripts", "models.R"))

# For DRY (Don't repeat yourself) parameters
source(here("Scripts", "DataAppendixScripts", "helpers.R"))

# For simulations caching
source(here("Scripts", "DataAppendixScripts", "cache_utils.R"))

# For loading existing simulations cache
source(here("Scripts", "DataAppendixScripts", "cache_loader.R"))

# Load existing simulations cache if available (cache_loader.R)
cache_load_existing()
```

## Power analysis with fixed N

The preceding sections determined the minimum $N$ required to achieve
adequate power across a range of conditions. This section takes a
different perspective: rather than searching for the smallest sufficient
sample size, it fixes $N = 350$ and asks how reliably that value
sustains the target power under the most demanding combined scenario —
10% MCAR missingness and mild non-normality simultaneously [@feng2023;
@simsem2021].

This design choice reflects a common applied situation in which the
sample size is constrained by practical or budgetary considerations.
Running a large number of replications at a fixed $N$ provides a direct
stability check: if power remains consistently above the target across
many random samples, the chosen sample size can be considered a robust
planning value for the full research design [@wang2021; @jobst2023].

```{r}
#| label: sim-fixed-popmodel

popmodel_fixed <- cache_or_run("popmodel_fixed", {
    sim(
        REP,                    # Number of replications (defined in helpers.R)
        n = 350,                # Fixed sample size
        pmMCAR = P_MCAR,        # Percentage of missing Completely At Random (MCAR) data (defined in helpers.R)
        analyzemodel,           # Model to analyze (defined in models.R)
        generate = popmodel,    # Population model to simulate from (defined in models.R)
        indDist = dist,         # Individual distribution of response variables (defined in helpers.R)
        std.lv = TRUE,          # Standardize latent variables to unit variance
        lavaanfun = "cfa",      # Use confirmatory factor analysis (CFA)
        estimator = "mlr",      # Use robust maximum likelihood (MLR) estimator
        missing = "fiml",       # Use Full Information Maximum Likelihood (FIML) for missing data
        seed = SEED,            # Random seed for reproducibility (defined in helpers.R)
        multicore = TRUE        # Use multiple cores for parallel processing
    )
})
```

```{r}
#| label: sim-fixed-h1model

h1model_fixed <- cache_or_run("h1model_fixed", {
    sim(
        REP,                    # Number of replications (defined in helpers.R)
        n = 350,                # Fixed sample size
        pmMCAR = P_MCAR,        # Percentage of missing Completely At Random (MCAR) data (defined in helpers.R)
        analyzemodel,           # Model to analyze (defined in models.R)
        generate = h1model,     # Population model with cross loadings (defined in models.R)
        indDist = dist,         # Individual distribution of response variables (defined in helpers.R)
        std.lv = TRUE,          # Standardize latent variables to unit variance
        lavaanfun = "cfa",      # Use confirmatory factor analysis (CFA)
        estimator = "mlr",      # Use robust maximum likelihood (MLR) estimator
        missing = "fiml",       # Use Full Information Maximum Likelihood (FIML) for missing data
        seed = SEED,            # Random seed for reproducibility (defined in helpers.R)
        multicore = TRUE        # Use multiple cores for parallel processing
    )
})
```

### Power curves for fit indices

With $N$ fixed, `plotPowerFit()` no longer traces power as a function
of sample size. Instead, the figure displays the empirical distributions
of each fit index under the reference and misspecified conditions across
replications, alongside the cutoff derived from the reference model at
the chosen $\alpha$ level. The visual separation between the two
distributions is the graphical counterpart of the power values reported
in the table below [@simsem2021; @mcneish2021].

```{r}
#| label: fig-power-curve-fixed
#| fig-cap: "Power curves for fit indices with fixed sample size."
#| fig-width: 8
#| fig-height: 5

plotPowerFit(
    h1model_fixed,    # Data simulated under H1 (with cross-loadings)
    popmodel_fixed,   # Data simulated under H0 (null model)
    alpha = ALPHA,    # Significance level for determining power (see helpers.R)
    usedFit = FITS    # RMSEA, SRMR, CFI, and TLI fit indices to display (see helpers.R)
)
```

### Power and cutoff values for fit indices

```{r}
#| label: tbl-power-cutoff-fixed
#| tbl-cap: "Power and fit index cutoff values with fixed sample size."

# Estimate power for rejecting the misspecified model (h1model_fixed)
# using fit-index cutoffs implied by the null model.
power_fixed <- getPowerFit(
  h1model_fixed,                 # Data simulated under H1 (with cross-loadings)
  nullObject = popmodel_fixed,   # Data simulated under H0 (null model)
  alpha = ALPHA,                 # Significance level (see helpers.R)
  usedFit = FITS                 # RMSEA, SRMR, CFI, and TLI fit indices to display (see helpers.R)
)

# Derive empirical cutoff values for the selected fit indices
# from the null-model sampling distribution under the same conditions.
cutoff_fixed <- getCutoff(
  popmodel_fixed,     # Data simulated under H0 (null model)
  alpha = ALPHA,      # Significance level (see helpers.R)
  usedFit = FITS      # RMSEA, SRMR, CFI, and TLI fit indices to display (see helpers.R)
)

# Subset power to the four requested fit indices
power_fixed_clean <- power_fixed[FITS]
cutoff_fixed_clean <- cutoff_fixed[1, FITS]

# Combine power and cutoff values into a compact summary table.
power_cutoff_fixed <- rbind(
  "Power"  = as.numeric(power_fixed_clean),
  "Cutoff" = as.numeric(cutoff_fixed_clean)
)

# Standardize column names for presentation.
colnames(power_cutoff_fixed) <- toupper(FITS)

# Print the table.
knitr::kable(power_cutoff_fixed, digits = 3, align = "c")
```

## Power analysis with traditional cutoffs

The same fixed-$N$ simulation is now re-evaluated using conventional
rule-of-thumb thresholds in place of the empirically derived cutoffs
used above. This is a deliberate contrast: by holding $N$, the
data-generating model, and the analysis model constant while changing
only the decision criterion, the comparison isolates the contribution
of the cutoff choice to the power conclusion [@mcneish2021;
@kyriazos2018].

```{r}
#| label: fig-curve-fixed-traditional
#| fig-cap: "Power curves using traditional fit index cutoffs with fixed sample size."
#| fig-width: 8
#| fig-height: 5

plotPowerFit(
    h1model_fixed,           # Data simulated under H1 (with cross-loadings)
    popmodel_fixed,          # Data simulated under H0 (null model)
    cutoff = RULE_OF_THUMB,  # Traditional fit-index cutoffs (see helpers.R)
    alpha = ALPHA,           # Significance level (see helpers.R)
    usedFit = FITS           # RMSEA, SRMR, CFI, and TLI fit indices to display (see helpers.R)
)
```

```{r}
#| label: tbl-power-cutoff-traditional-fixed
#| tbl-cap: "Power for traditional fit index cutoffs with fixed sample size."

# Estimate power using traditional fit-index cutoffs.
power_traditional <- getPowerFit(
  h1model_fixed,     # Data simulated under H1 (with cross-loadings)
  RULE_OF_THUMB,     # Traditional fit-index cutoffs (see helpers.R)
  alpha = ALPHA,     # Significance level (see helpers.R)
  usedFit = FITS     # RMSEA, SRMR, CFI, and TLI fit indices to display (see helpers.R)
)

# Convert to a one-row table with column headers.
power_traditional_tbl <- t(as.data.frame(power_traditional))

# Standardize column names for presentation.
colnames(power_traditional_tbl) <- toupper(names(power_traditional))
rownames(power_traditional_tbl) <- "Power"

# Print the table.
knitr::kable(power_traditional_tbl, digits = 3, align = "c")
```


