---
title: "Simulation with Fixed Sample Size"
execute:
  echo: false
  warning: false
  message: false
---

## Setup

```{r}
#| label: setup
#| include: false

library(here)
source(here("Scripts", "DataAppendixScripts", "models.R"))
source(here("Scripts", "DataAppendixScripts", "helpers.R"))
source(here("Scripts", "DataAppendixScripts", "cache_utils.R"))
source(here("Scripts", "DataAppendixScripts", "cache_loader.R"))

cache_load_existing()
```


## Power Analysis with fixed N

```{r}
#| label: sim-fixed-popmodel
#| cache: true

popmodel_fixed <- cache_or_run("popmodel_fixed", {
    sim(
        REP,
        n = 350,
        pmMCAR = 0.10,
        analyzemodel,
        generate = popmodel,
        indDist = dist,       # See the `dist` object defined in helpers.R for the range of skewness and kurtosis values.
        std.lv = TRUE,
        lavaanfun = "cfa",
        estimator = "mlr",
        missing = "fiml",
        seed = SEED,
        multicore = TRUE
    )
})
```


```{r}
#| label: sim-fixed-h1model
#| cache: true

h1model_fixed <- cache_or_run("h1model_fixed", {
    sim(
        REP,
        n = 350,
        pmMCAR = 0.10,
        analyzemodel,
        generate = h1model,
        indDist = dist,      # See the `dist` object defined in helpers.R for the range of skewness and kurtosis values.
        std.lv = TRUE,
        lavaanfun = "cfa",
        estimator = "mlr",
        missing = "fiml",
        seed = SEED,
        multicore = TRUE
    )
})
```

### Power curves for fit indices

This chunk evaluates the main power analysis under a **fixed** sample size, $N = 350$. Instead of searching for the minimum required $N$, the simulation now examines how stable this preselected sample size is for rejecting a substantively relevant misspecified model across repeated samples.

Data are generated from the population model (`popmodel`), while the fitted analysis contrasts the intended analysis model against the misspecified comparison model (`h1model`), which adds three cross-loadings. In this way, the chunk estimates how often the chosen fit indices can distinguish the target measurement structure from a plausible but inferior alternative under repeated Monte Carlo replications [@moshagen2024; @feng2023].

The main outputs summarize this discrimination from complementary angles. `plotPowerFit(logistic = FALSE)` shows the empirical distributions of the fit indices under the competing data-generating conditions and the cutoff used to separate them. `getPowerFit` reports the achieved power at the fixed $N$, and `getCutoff` reports the empirical cutoffs derived from the reference model [@simsem2021].

```{r}
#| label: fig-power-curve-fixed
#| fig-cap: "Power curves for fit indices with fixed sample size."
#| fig-width: 8
#| fig-height: 5

plotPowerFit(
    h1model_fixed,
    popmodel_fixed,
    alpha = ALPHA,
    usedFit = FITS
)
```

### Power cutoff values for fit indices

```{r}
#| label: tbl-power-cutoff-fixed
#| tbl-cap: "Power and fit index cutoff values with fixed sample size."

# Estimate power for rejecting the misspecified model (h1model_fixed)
# using fit-index cutoffs implied by the null model.
power_fixed <- getPowerFit(
  h1model_fixed,
  nullObject = popmodel_fixed,
  alpha = ALPHA,
  usedFit = FITS
)

# Derive empirical cutoff values for the selected fit indices
# from the null-model sampling distribution under the same conditions.
cutoff_fixed <- getCutoff(
  popmodel_fixed,
  alpha = ALPHA,
  usedFit = FITS
)

# Combine power and cutoff values into a compact summary table.
power_cutoff_fixed <- rbind(
  "Power" = as.numeric(power_fixed),
  "Cutoff" = as.numeric(cutoff_fixed[1, ])
)

colnames(power_cutoff_fixed) <- toupper(names(power_fixed))

knitr::kable(power_cutoff_fixed, digits = 3, align = "c")
```

## Power analysis with traditional fits

In the following chunk, the same fixed-$N$ design is re-evaluated using traditional rule-of-thumb cutoffs. This is included only as an illustration of how generic thresholds may lead to very different conclusions from empirically calibrated cutoffs tailored to the specific model, misspecification, and sample size [@jak2021; @feng2023].

```{r}
#| label: fig-curve-fixed-traditional
#| fig-cap: "Power curves using traditional fit index cutoffs with fixed sample size."
#| fig-width: 8
#| fig-height: 5

plotPowerFit(
    h1model_fixed,
    popmodel_fixed,
    cutoff = RULE_OF_THUMB,
    alpha = ALPHA,
    usedFit = FITS
)
```


```{r}
#| label: tbl-power-cutoff-traditional
#| tbl-cap: "Power for traditional fit index cutoffs with fixed sample size."

# Estimate power using traditional fit-index cutoffs.
power_traditional <- getPowerFit(
  h1model_fixed,
  RULE_OF_THUMB,
  alpha = ALPHA,
  usedFit = FITS
)

# Convert to a one-row table with column headers.
power_traditional_tbl <- t(as.data.frame(power_traditional))

colnames(power_traditional_tbl) <- toupper(names(power_traditional))
rownames(power_traditional_tbl) <- "Power"

knitr::kable(power_traditional_tbl, digits = 3, align = "c")
```


