---
title: "Simulation with Fixed Sample Size"
execute:
  echo: false
  warning: false
  message: false
---

## Setup

```{r}
#| label: setup
#| include: false

library(here)
source(here("Scripts", "DataAppendixScripts", "models.R"))
source(here("Scripts", "DataAppendixScripts", "helpers.R"))
source(here("Scripts", "DataAppendixScripts", "cache_utils.R"))
source(here("Scripts", "DataAppendixScripts", "cache_loader.R"))

cache_load_existing()
```


## Power Analysis with fixed N

Ideial condition: PA it not too sensitive to distributional assumptions, missing data, etc. So, we use ideal for time consuming simulations. We can also do the same for the robust conditions, but it will take more time.

```{r}
#| label: sim-fixed-popmodel
#| cache: true

popmodel_fixed <- cache_or_run("popmodel_fixed", {
    sim(
        REP,
        n = 350,
        pmMCAR = 0.10,
        analyzemodel,
        generate = popmodel,
        indDist = dist,       # See the `dist` object defined in helpers.R for the range of skewness and kurtosis values.
        std.lv = TRUE,
        lavaanfun = "cfa",
        estimator = "mlr",
        missing = "fiml",
        seed = SEED,
        multicore = TRUE
    )
})
```


```{r}
#| label: sim-fixed-h1model
#| cache: true

h1model_fixed <- cache_or_run("h1model_fixed", {
    sim(
        REP,
        n = 350,
        pmMCAR = 0.10,
        analyzemodel,
        generate = h1model,
        indDist = dist,      # See the `dist` object defined in helpers.R for the range of skewness and kurtosis values.
        std.lv = TRUE,
        lavaanfun = "cfa",
        estimator = "mlr",
        missing = "fiml",
        seed = SEED,
        multicore = TRUE
    )
})
```

### Power curves for fit indices

```{r}
#| label: fig-power-curve-fixed
#| fig-cap: "Power curves for fit indices with fixed sample size."
#| fig-width: 8
#| fig-height: 5

plotPowerFit(
    h1model_fixed,
    popmodel_fixed,
    alpha = ALPHA,
    usedFit = FITS
)
```

### Power cutoff values for fit indices

You extract the cutoff values for fit indices at a given alpha level and sample size to determine the threshold for acceptable fit under the null model. This allows you to see how stringent the fit criteria are at different sample sizes.

```{r}
#| label: tbl-power-cutoff-fixed
#| tbl-cap: "Power and fit index cutoff values with fixed sample size."

# Estimate power for rejecting the misspecified model (h1model_fixed)
# using fit-index cutoffs implied by the null model.
power_fixed <- getPowerFit(
  h1model_fixed,
  nullObject = popmodel_fixed,
  alpha = ALPHA,
  usedFit = FITS
)

# Derive empirical cutoff values for the selected fit indices
# from the null-model sampling distribution under the same conditions.
cutoff_fixed <- getCutoff(
  popmodel_fixed,
  alpha = ALPHA,
  usedFit = FITS
)

# Combine power and cutoff values into a compact summary table.
power_cutoff_fixed <- rbind(
  "Power" = as.numeric(power_fixed),
  "Cutoff" = as.numeric(cutoff_fixed[1, ])
)

colnames(power_cutoff_fixed) <- toupper(names(power_fixed))

knitr::kable(power_cutoff_fixed, digits = 3, align = "c")
```

## Power analysis with traditional fits

```{r}
#| label: fig-fit-curve-fixed
#| fig-cap: "Power curves using traditional fit index cutoffs with fixed sample size."
#| fig-width: 8
#| fig-height: 5

plotPowerFit(
    h1model_fixed,
    popmodel_fixed,
    cutoff = RULE_OF_THUMB,
    alpha = ALPHA,
    usedFit = FITS
)
```


```{r}
#| label: tbl-power-cutoff-traditional
#| tbl-cap: "Power for traditional fit index cutoffs with fixed sample size."

# Estimate power using traditional fit-index cutoffs.
power_traditional <- getPowerFit(
  h1model_fixed,
  RULE_OF_THUMB,
  alpha = ALPHA,
  usedFit = FITS
)

# Convert to a one-row table with column headers.
power_traditional_tbl <- t(as.data.frame(power_traditional))

colnames(power_traditional_tbl) <- toupper(names(power_traditional))
rownames(power_traditional_tbl) <- "Power"

knitr::kable(power_traditional_tbl, digits = 3, align = "c")
```


