---
title: "Post Hoc Power Analysis"
execute:
  echo: false
  warning: false
  message: false
---

## Setup

```{r}
#| label: setup
#| include: false

library(here)
source(here("Scripts", "DataAppendixScripts", "models.R"))
source(here("Scripts", "DataAppendixScripts", "helpers.R"))
source(here("Scripts", "DataAppendixScripts", "cache_utils.R"))
source(here("Scripts", "DataAppendixScripts", "cache_loader.R"))

cache_load_existing()

```

The post hoc analysis is based on the real dataset from @rogers2022,
publicly available in Mendeley Data. The data contain WHOQOL-BREF
responses from 1,047 participants and are loaded directly from the
repository URL. The sample size is extracted from the object and
reused throughout this section, and the model degrees of freedom are
computed from `analyzemodel` as a required input for the analytical
functions.

```{r}
#| label: load-realdata

url_data <- "https://data.mendeley.com/datasets/rdky78bk8r/2/files/b58a7054-4978-4239-81db-00f9efd86e44/WHOQOL_Data.dat"
realdata <- read.table(url_data)

url_labels <- "https://data.mendeley.com/datasets/rdky78bk8r/2/files/ae2a5c01-2fc2-43e5-b31c-8db8cc5bff8a/WHOQOL_Labels.txt"
column_labels <- read.table(url_labels)
column_labels <- column_labels$V1
names(realdata) <- gsub("_[FPSA]$", "", column_labels)

n_real <- nrow(realdata) # Sample size of the real data

df_model <- 245 # Degrees of freedom with std.lv = TRUE
```

## Post hoc power analysis

### With simulation

Post hoc simulation-based power analysis follows four steps: fit the
analysis model to the real data to obtain the empirical parameter
estimates; simulate data from that fitted model to construct the null
reference distribution; simulate data from the misspecified model to
construct the alternative distribution; and evaluate how well each fit
index separates the two distributions at the observed $N$
[@simsem2021; @wang2021].

Grounding the reference distribution in the fitted model rather than in
the theoretical population model is what distinguishes this from the a
priori analyses above: here, the population structure is implied by the
data rather than specified in advance, which makes the power estimate
reflect the evidence actually available in the collected sample
[@feng2023; @jobst2023].

```{r}
#| label: realdata-model

fit_real <- cfa(
  analyzemodel,
  data      = realdata,
  estimator = "mlr",
  std.lv    = TRUE
)
```

```{r}
#| label: sim-null-posthoc

realmodel_null <- cache_or_run("realmodel_null", {
  sim(
      nRep      = REP,        # See the `REP` object defined in helpers.R for the number of replications used in this scenario.
      n         = n_real,
      analyzemodel,
      generate  = fit_real,
      std.lv    = TRUE,
      lavaanfun = "cfa",
      estimator = "mlr",
      seed      = SEED,
      multicore = TRUE
  )
})
```

```{r}
#| label: sim-alt-posthoc

h1model_alt <- cache_or_run("h1model_alt", {
  sim(
      nRep      = REP,        # See the `REP` object defined in helpers.R for the number of replications used in this scenario.
      n         = n_real,
      analyzemodel,
      generate  = h1model,
      std.lv    = TRUE,
      lavaanfun = "cfa",
      estimator = "mlr",
      seed      = SEED,
      multicore = TRUE
  )
})
```

The figure below displays the fit-index distributions for the
alternative and null conditions, alongside the empirical cutoff derived
from the reference model at $\alpha$. The degree of visual separation
between the two distributions provides an intuitive picture of whether
the observed $N$ is sufficient to support reliable model discrimination
[@simsem2021; @mcneish2021].

```{r}
#| label: fig-posthoc-plotpower
#| fig-cap: "Power curve for post hoc analysis based on simulation."
#| fig-width: 8
#| fig-height: 5

# Plot the fit-index distributions for the alternative model and compare them
# with cutoffs obtained from the null model at level alpha.
# In simsem, this visualizes how well the selected global fit indices separate
# data generated under the null and alternative conditions.

plotPowerFit(
    h1model_alt,
    nullObject = realmodel_null,
    alpha      = ALPHA,
    usedFit    = FITS
)
```

The table below reports the corresponding power values for each index:
the proportion of alternative-condition replications that fall beyond
the null-model cutoff at the chosen $\alpha$ level [@simsem2021;
@wang2023].

```{r}
#| label: tbl-posthoc-power
#| tbl-cap: "Power for post hoc analysis based on simulation."

# Return post hoc power for each selected fit index.
# The cutoffs are internally derived from the null-object distribution,
# so the table summarizes the empirical rejection rate of the alternative model
# under the chosen set of fit criteria.

posthoc_pow <- getPowerFit(
    h1model_alt,
    nullObject = realmodel_null,
    alpha      = ALPHA,
    usedFit    = FITS
)

# Subset power to the four requested fit indices (FITS: rmsea, srmr, cfi, tli)
# ensuring TRADITIONALCHI is excluded.
posthoc_pow_clean <- posthoc_pow[FITS]

knitr::kable(
  as.data.frame(t(as.numeric(posthoc_pow_clean))),
  col.names = toupper(FITS),
  digits = 3,
  align = "c"
)
```

### With semPower

For comparison, the same post hoc question is addressed analytically
through three specifications of increasing specificity. The first is
model-free and sets the target discrepancy at $\text{RMSEA} = 0.05$;
the second is model-based and derives the effect from the discrepancy
between the misspecified population model (`h1model`) and the empirically fitted model; the
third specifies an explicit nested-model comparison by supplying both a
null and an alternative model, so the effect reflects a targeted
structural difference rather than global misfit against the saturated
model [@moshagen2024; @jak2021].

```{r}
#| label: fig-posthoc-rmsea
#| fig-cap: "Post hoc power analysis for RMSEA = 0.05."
#| fig-width: 8
#| fig-height: 5

# Analytical post hoc power for a global fit test.
# The target effect is fixed directly as RMSEA = 0.05.
# H0 is exact fit, alpha is 0.05, and the sample size is 1,047.

ph_rmsea <- semPower::semPower.postHoc(
  effect         = 0.05,
  effect.measure = "RMSEA",  # Target discrepancy expressed as RMSEA
  alpha          = ALPHA,
  N              = n_real,
  df             = df_model,
  p              = P         # Only relevant for indices such as GFI and AGFI
)

summary(ph_rmsea)
```

In the model-based specification below, `semPower.powerLav()` fits the
analysis model to the population-implied covariance matrix and translates
the resulting discrepancy into the noncentrality parameter. Because
`modelH1` is omitted, the comparison is against the saturated model,
making this a global test of fit rather than a test of a specific
restriction [@moshagen2024; @jobst2023].

```{r}
#| label: fig-posthoc-realmodel
#| fig-cap: "Post hoc power analysis for the real model."
#| fig-width: 8
#| fig-height: 5

# Analytical post hoc power based on lavaan models.
# modelPop defines the population structure (h1model with cross-loadings).
# modelH0 is the focal model to be tested.
# Because modelH1 is omitted, the comparison is against the saturated model.

ph_realmodel <- semPower::semPower.powerLav(
  type       = "post-hoc",
  alpha      = ALPHA,
  N          = n_real,
  modelPop   = h1model,
  modelH0    = fit_real,
  lavOptions = list(std.lv = TRUE)
)

summary(ph_realmodel)
```


