---
title: "Post Hoc Power Analysis"
execute:
  echo: false
  warning: false
  message: false
---

## Setup

```{r}
#| label: setup
#| include: false

# For defining file paths
library(here)

# For CFA population models
source(here("Scripts", "DataAppendixScripts", "models.R"))

# For DRY (Don't repeat yourself) parameters
source(here("Scripts", "DataAppendixScripts", "helpers.R"))

# For simulations caching
source(here("Scripts", "DataAppendixScripts", "cache_utils.R"))

# For loading existing simulations cache
source(here("Scripts", "DataAppendixScripts", "cache_loader.R"))

# Load existing simulations cache if available (cache_loader.R)
cache_load_existing()
```

The post hoc analysis is based on the real dataset from @rogers2022,
publicly available in Mendeley Data. The data contain WHOQOL-BREF
responses from 1,047 participants and are loaded directly from the
repository URL. The sample size is extracted from the object and
reused throughout this section, and the model degrees of freedom are
computed from `analyzemodel` as a required input for the analytical
functions.

```{r}
#| label: load-realdata

# Load the real dataset from Mendeley Data
url_data <- "https://data.mendeley.com/datasets/rdky78bk8r/2/files/b58a7054-4978-4239-81db-00f9efd86e44/WHOQOL_Data.dat"
realdata <- read.table(url_data)

# Load the corresponding labels for the real dataset
url_labels <- "https://data.mendeley.com/datasets/rdky78bk8r/2/files/ae2a5c01-2fc2-43e5-b31c-8db8cc5bff8a/WHOQOL_Labels.txt"
column_labels <- read.table(url_labels)

# Assign the column labels to the real dataset
column_labels <- column_labels$V1
names(realdata) <- gsub("_[FPSA]$", "", column_labels)

# Extract the sample size from the real dataset
n_real <- nrow(realdata)

# Compute the model degrees of freedom from `analyzemodel`
df_model <- semPower.getDf(analyzemodel)-4 # See `02_analytical_power.qmd`
```

## Post hoc power analysis

### With simulation

Post hoc simulation-based power analysis follows four steps: fit the
analysis model to the real data to obtain the empirical parameter
estimates; simulate data from that fitted model to construct the null
reference distribution; simulate data from the misspecified model to
construct the alternative distribution; and evaluate how well each fit
index separates the two distributions at the observed $N$
[@simsem2021; @wang2021].

Grounding the reference distribution in the fitted model rather than in
the theoretical population model is what distinguishes this from the a
priori analyses above: here, the population structure is implied by the
data rather than specified in advance, which makes the power estimate
reflect the evidence actually available in the collected sample
[@feng2023; @jobst2023].

```{r}
#| label: realdata-model

fit_real <- cfa(
  analyzemodel,           # Model to analyze (see `models.R`)
  data      = realdata,   # Real data from Mendeley Data
  estimator = "mlr",      # Robust maximum likelihood estimator
  std.lv    = TRUE        # Standardized latent variables
)
```

```{r}
#| label: sim-null-posthoc

realmodel_null <- cache_or_run("realmodel_null", {
  sim(
      nRep      = REP,        # Number of replications (defined in helpers.R)
      n         = n_real,     # Sample size of the real data
      analyzemodel,           # Model to analyze (see `models.R`)
      generate  = fit_real,   # Fitted model from real data
      std.lv    = TRUE,       # Standardized latent variables
      lavaanfun = "cfa",      # Use confirmatory factor analysis (CFA) function
      estimator = "mlr",      # Robust maximum likelihood estimator
      seed      = SEED,       # Random seed for reproducibility (defined in helpers.R) 
      multicore = TRUE        # Use multiple cores for parallel processing
  )
})
```

```{r}
#| label: sim-alt-posthoc

h1model_alt <- cache_or_run("h1model_alt", {
  sim(
      nRep      = REP,        # Number of replications (defined in helpers.R)
      n         = n_real,     # Sample size of the real data
      analyzemodel,           # Model to analyze (see `models.R`)
      generate  = h1model,    # Population model with misspecification (see `models.R`)
      std.lv    = TRUE,       # Standardized latent variables
      lavaanfun = "cfa",      # Use confirmatory factor analysis (CFA) function
      estimator = "mlr",      # Robust maximum likelihood estimator
      seed      = SEED,       # Random seed for reproducibility (defined in helpers.R) 
      multicore = TRUE        # Use multiple cores for parallel processing
  )
})
```

The figure below displays the fit-index distributions for the
alternative and null conditions, alongside the empirical cutoff derived
from the reference model at $\alpha$. The degree of visual separation
between the two distributions provides an intuitive picture of whether
the observed $N$ is sufficient to support reliable model discrimination
[@simsem2021; @mcneish2021].

```{r}
#| label: fig-posthoc-plotpower
#| fig-cap: "Power curve for post hoc analysis based on simulation."
#| fig-width: 8
#| fig-height: 5

plotPowerFit(
    h1model_alt,                  # Data simulated under H1 (with cross-loadings)
    nullObject = realmodel_null,  # Data simulated under H0 (no cross-loadings)
    alpha      = ALPHA,           # Significance level (see helpers.R)
    usedFit    = FITS             # RMSEA, SRMR, CFI, and TLI fit indices to display (see helpers.R)
)
```

The table below reports the corresponding power values for each index:
the proportion of alternative-condition replications that fall beyond
the null-model cutoff at the chosen $\alpha$ level [@simsem2021;
@wang2023].

```{r}
#| label: tbl-posthoc-power
#| tbl-cap: "Power for post hoc analysis based on simulation."

# Estimate power for rejecting the misspecified model (h1model_alt)
# using fit-index cutoffs implied by the null model.
posthoc_pow <- getPowerFit(
    h1model_alt,                 # Data simulated under H1 (with cross-loadings)
    nullObject = realmodel_null, # Data simulated under H0 (no cross-loadings)
    alpha      = ALPHA,          # Significance level (see helpers.R)
    usedFit    = FITS            # RMSEA, SRMR, CFI, and TLI fit indices to display (see helpers.R)
)

# Subset power to the four requested fit indices
posthoc_pow_clean <- posthoc_pow[FITS]

# Return power values as a table
knitr::kable(
  as.data.frame(t(as.numeric(posthoc_pow_clean))),
  col.names = toupper(FITS),
  digits = 3,
  align = "c"
)
```

### With semPower

For comparison, the same post hoc question is addressed analytically
through two specifications. The first is model-free and sets the target discrepancy at $\text{RMSEA} = 0.05$;
the second is model-based and derives the effect from the discrepancy
between the misspecified population model (`h1model`) and the empirically fitted model [@moshagen2024; @jak2021].

```{r}
#| label: fig-posthoc-rmsea
#| fig-cap: "Post hoc power analysis for RMSEA = 0.05."
#| fig-width: 8
#| fig-height: 5

# Analytical post hoc power for a global fit test.
# The target effect is fixed directly as RMSEA = 0.05.

ph_rmsea <- semPower::semPower.postHoc(
  effect         = 0.05,        # RMSEA under H1
  effect.measure = "RMSEA",     # Target discrepancy expressed as RMSEA
  alpha          = ALPHA,       # Significance level (see helpers.R)
  N              = n_real,      # Sample size from real data
  df             = df_model,    # Degrees of freedom of the model
  p              = P            # Only relevant for indices such as GFI and AGFI
)

# Summary of the power analysis
summary(ph_rmsea)
```

In the model-based specification below, `semPower.powerLav()` fits the
analysis model to the population-implied covariance matrix and translates
the resulting discrepancy into the noncentrality parameter. Because
`modelH1` is omitted, the comparison is against the saturated model,
making this a global test of fit rather than a test of a specific
restriction [@moshagen2024; @jobst2023].

```{r}
#| label: fig-posthoc-realmodel
#| fig-cap: "Post hoc power analysis for the real model."
#| fig-width: 8
#| fig-height: 5

# Analytical post hoc power based on lavaan models.

ph_realmodel <- semPower::semPower.powerLav(
  type       = "post-hoc",          # Post hoc power analysis
  alpha      = ALPHA,               # Significance level (see helpers.R)
  N          = n_real,              # Sample size from real data
  modelPop   = h1model,             # Population model with cross loadings (see `models.R`)
  modelH0    = fit_real,            # Fitted model from real data
  lavOptions = list(std.lv = TRUE)  # Standardized latent variables
)

# Summary of the power analysis
summary(ph_realmodel)
```


